### 2022-05-28 paper/paper.tex
Add paper template

% My template
\documentclass[12pt, letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[amsthm]{ntheorem}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{url}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    citecolor=black,
    filecolor=black,
    linkcolor=black,
    urlcolor=black
}
\widowpenalty10000
\clubpenalty10000
\setlength{\parskip}{\baselineskip}
\setlength{\parindent}{0pt}
\setlist[enumerate]{topsep=0pt}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{corollary}{Corollary}[theorem]
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}\newtheorem*{definition}{Definition}
\theoremstyle{definition}\newtheorem*{example}{Example}
\theoremstyle{definition}\newtheorem*{remark}{Remark}
\newcommand{\vect}[1]{\boldsymbol{#1}}
\newcommand{\mb}[1]{\mathbf{#1}}
\def\RR{\mathbb{R}}
\def\ZZ{\mathbb{Z}}
\def\EE{\mathbb{E}}
\def\PP{\mathbb{P}}
\def\QQ{\mathbb{Q}}
\def\II{\mathbb{I}}
\def\CC{\mathbb{C}}
\def\NN{\mathbb{N}}
\def\OO{\mathbb{O}}
\def\Ll{\mathcal{L}}
\def\Mm{\mathcal{M}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}

\title{\vspace{-2.0cm}Mitigating the Effect of Sampling Bias on Recommender Systems through Quantization}
\author{Fengyu Li, Sarah Dean}
\date{}

\begin{document}
\maketitle

\begin{abstract}
    Placleholder.
\end{abstract}

\section{Introduction}

\section{Observations}

\end{document}

### 2022-06-08 data_utils.py
Fix a bug that causes dimension problems

@@ -5,6 +5,7 @@ from random import random
 from math import exp
 import reclab
 
+EPSILON = 1e-8
 
 def generate_ground_truth_matrix(dimensions, environment='random'):
     """
@@ -25,7 +26,7 @@ def generate_ground_truth_matrix(dimensions, environment='random'):
     else:
         env = reclab.make(environment, num_users=m, num_items=n, noise=0)
         env.reset()
-        return env._get_dense_ratings()
+        return np.array(env._get_dense_ratings())
 
 
 def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=None, beta=1, noise=0.05):
@@ -81,7 +82,9 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
         softmax_result = exps / np.sum(exps)
         # Scale mean to one
         softmax_result /= np.mean(softmax_result)
-        P = np.tile(softmax_result, m)
+        P = np.tile(softmax_result, (m, 1))
+        assert abs(P.mean()) - 1 < EPSILON
+        assert P.shape == matrix.shape
 
         ratings = {}
         for i in range(m):
@@ -96,7 +99,9 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
         softmax_result = exps / np.sum(exps)
         # Scale mean to one
         softmax_result /= np.mean(softmax_result)
-        P = np.tile(softmax_result, (n, 1))
+        P = np.tile(softmax_result.reshape(1, m), (n, 1))
+        assert abs(P.mean()) - 1 < EPSILON
+        assert P.shape == matrix.shape
 
         ratings = {}
         for i in range(m):
@@ -109,7 +114,7 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
         P = np.exp(matrix)
         P /= np.sum(P)
         P /= np.mean(P)
-        print(f'Mean: {P.mean()}')
+        assert abs(P.mean()) - 1 < EPSILON
     
         ratings = {}
         for i in range(m):
@@ -147,12 +152,7 @@ def sample(value, sample_prob=0.1):
     else:
         return value
 
-
 if __name__ == '__main__':
-    truth = generate_ground_truth_matrix(
-        (1000, 1000), environment='latent-static-v1')
-    # print(truth)
[…]

### 2022-06-09 data_utils.py
Edit during commit

@@ -111,7 +111,7 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
         return users, items, ratings
 
     elif bias == 'full underlying':
-        P = np.exp(matrix)
+        P = np.exp(beta * matrix)
         P /= np.sum(P)
         P /= np.mean(P)
         assert abs(P.mean()) - 1 < EPSILON
@@ -156,3 +156,9 @@ if __name__ == '__main__':
     truth = generate_ground_truth_matrix((1000, 1000), environment='latent-dynamic-v1')
     assert truth.shape == (1000, 1000)
     users, items, ratings = ground_truth_matrix_to_dataset(truth, quantization='onetofive', bias='active user')
+
+    count = 0
+    for rating in ratings.values():
+        if rating is not None:
+            count += 1
+    print(count / (1000 * 1000))        

### 2022-06-10 data_utils.py
Complete doc

@@ -36,13 +36,19 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
     Bias can be introduced in the dataset.
 
     Parameters:
-        R: a mxn matrix that contains the ground truth for each pair of m users and n items.
-        sample_prob: the probability of sampling a chosen pair of users and items.
+        matrix: a mxn matrix that contains the ground truth for each pair of m users and n items.
+        quantization: the way to quantize / discretize the ratings.
+        sample_prob: the expected probability of sampling a chosen pair of users and items.
         bias: the type of sampling bias.
-        shuffle: whether to shuffle the dataset by users.
+        beta: hyperparameter that controls the degress of the bias.
+        noise: hyperparameter that controls the degree of noise.
 
     Returns:
-        Same as reclab.data_utils.read_dataset()
+        users: a dictionary whose ith entry is a list of observed ratings for the ith user.
+        itmes: a dictionary whose ith entry is a list of observed ratings for the ith item.
+        ratings: a dictionary whose entry with key (i, j) is the observed rating for user i on item j.
+        P: P[i, j] is the probability of observing rating i on item j.
+        R: R is a quantized version of the input matrix. Noise is also added.
     """
     m, n = matrix.shape
     np.random.shuffle(matrix)
@@ -74,54 +80,57 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
             for j in range(n):
                 ratings[(i, j)] = sample(R[i, j], sample_prob)
         users, items = generate_users_items(ratings, m, n)
-        return users, items, ratings
+        return users, items, ratings, P, R
 
     elif bias == 'popularity':
         average_ratings = np.mean(matrix, axis=0)
         exps = np.exp(beta * average_ratings)
         softmax_result = exps / np.sum(exps)
-        # Scale mean to one
+        # Scale mean to sample_prob
         softmax_result /= np.mean(softmax_result)
+        softmax_result *= sample_prob
         P = np.tile(softmax_result, (m, 1))
-        assert abs(P.mean()) - 1 < EPSILON
+        assert abs(P.mean()) - sample_prob < EPSILON
         assert P.shape == matrix.shape
 
         ratings = {}
         for i in range(m):
             for j in range(n):
[…]

### 2022-06-10 data_utils.py
Add function to calculate correlation

@@ -1,11 +1,11 @@
-from multiprocessing.sharedctypes import Value
 import numpy as np
-import pandas as pd
+from scipy.stats import pearsonr, spearmanr
 from random import random
 from math import exp
 import reclab
 
 EPSILON = 1e-8
+
 
 def generate_ground_truth_matrix(dimensions, environment='random'):
     """
@@ -124,7 +124,7 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
         P /= np.mean(P)
         P *= sample_prob
         assert abs(P.mean()) - sample_prob < EPSILON
-    
+
         ratings = {}
         for i in range(m):
             for j in range(n):
@@ -133,7 +133,7 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
         return users, items, ratings, P, R
 
     else:
-        raise ValueError('Bias method not supported.')    
+        raise ValueError('Bias method not supported.')
 
 
 def generate_users_items(ratings, m, n):
@@ -161,13 +161,28 @@ def sample(value, sample_prob=0.1):
     else:
         return value
 
-if __name__ == '__main__':
-    truth = generate_ground_truth_matrix((1000, 1000), environment='latent-dynamic-v1')
-    assert truth.shape == (1000, 1000)
-    users, items, ratings, P, R = ground_truth_matrix_to_dataset(truth, quantization='onetofive', bias='active user')
 
-    count = 0
-    for rating in ratings.values():
-        if rating is not None:
-            count += 1
-    print(count / (1000 * 1000))        
+def correlation(P, matrix, correlation='pearson'):
+    """
+    Helper function that computes the correlation between two matrices.
+    """
+    if correlation == 'pearson':
+        return pearsonr(P.flatten(), matrix.flatten())[0]
+    elif correlation == 'spearman':
+        return spearmanr(P.flatten(), matrix.flatten())[0]
+
+
+if __name__ == '__main__':
+    truth = generate_ground_truth_matrix(
+        (1000, 1000), environment='latent-dynamic-v1')
+    assert truth.shape == (1000, 1000)
+    users, items, ratings, P, R = ground_truth_matrix_to_dataset(
+        truth, quantization='onetofive', bias='popularity')
+
+    # count = 0
+    # for rating in ratings.values():
+    #     if rating is not None:
+    #         count += 1
+    # print(count / (1000 * 1000))
+
+    print(correlation(P, truth))

### 2022-06-11 data_utils.py
Plot beta vs correlation

@@ -72,8 +72,6 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
     else:
         raise ValueError('Quantization scale not supported.')
 
-    print(f'Ground truth after normalization and quantization: {R}')
-
     if bias is None:
         ratings = {}
         for i in range(m):
@@ -175,14 +173,7 @@ def correlation(P, matrix, correlation='pearson'):
 if __name__ == '__main__':
     truth = generate_ground_truth_matrix(
         (1000, 1000), environment='latent-dynamic-v1')
-    assert truth.shape == (1000, 1000)
     users, items, ratings, P, R = ground_truth_matrix_to_dataset(
         truth, quantization='onetofive', bias='popularity')
 
-    # count = 0
-    # for rating in ratings.values():
-    #     if rating is not None:
-    #         count += 1
-    # print(count / (1000 * 1000))
-
-    print(correlation(P, truth))
+    print(correlation(P, truth, correlation='spearman'))

### 2022-06-11 data_utils.py
Plot again

@@ -1,5 +1,5 @@
 import numpy as np
-from scipy.stats import pearsonr, spearmanr
+from scipy.stats import pearsonr, spearmanr, kendalltau
 from random import random
 from math import exp
 import reclab
@@ -168,6 +168,8 @@ def correlation(P, matrix, correlation='pearson'):
         return pearsonr(P.flatten(), matrix.flatten())[0]
     elif correlation == 'spearman':
         return spearmanr(P.flatten(), matrix.flatten())[0]
+    elif correlation == 'kendalltau':
+        return kendalltau(P.flatten(), matrix.flatten())[0]
 
 
 if __name__ == '__main__':

### 2022-06-11 data_utils.py
Remove garbage files

@@ -23,6 +23,10 @@ def generate_ground_truth_matrix(dimensions, environment='random'):
 
     if environment == 'random':
         return np.random.rand(dimensions)
+    elif environment == 'ml-100k-v1':
+        env = reclab.make(environment)
+        env.reset()
+        return np.array(env._get_dense_ratings())
     else:
         env = reclab.make(environment, num_users=m, num_items=n, noise=0)
         env.reset()

### 2022-06-14 paper/paper.tex
Edit template
Add some draft

@@
 \def\OO{\mathbb{O}}
+\def\HH{\mathbb{H}}
 \def\Ll{\mathcal{L}}
@@
 
-\section{Observations}
+\section{Draft Area}
+ (Consider using 2-dimensional Gaussian?)\\
+Assume the (flattened) ground truth ratings is a standard Gaussian with differential entropy
+\begin{align*} \HH(R\sim N(0, 1)) & = -\EE[\log R]                    \\ & = -\int_{-\infty}^\infty \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}} \log \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}} \, dx \\
+                                  & = \frac{1}{2}\log 2\pi e          \\
+                                  & = 2.05
+\end{align*}
+We can show that quantization (discretization) loses information. Consider
+a simple binary quantization that splits the ratings around the mean. Denote the quantized r.v. as $Q_1$.
+\[ \HH(Q_1) = -2 \cdot \frac{1}{2} \log \frac{1}{2} = 1 \]
+Consider another quantized r.v. $Q_2$ that divides the distribution four-fold based on the four quantiles.
+\[ \HH(Q_2) = -4 \cdot \frac{1}{4} \log \frac{1}{4} = 2 \]
+We can see that different quantization schemes loses different amount of information.
+\[ \HH(R) > \HH(Q_2) > \HH(Q_1) \]
 
+Next: discussion
 \end{document}

### 2022-06-14 data_utils.py

@@ -178,8 +178,8 @@ def correlation(P, matrix, correlation='pearson'):
 
 if __name__ == '__main__':
     truth = generate_ground_truth_matrix(
-        (1000, 1000), environment='latent-dynamic-v1')
+        (1000, 1000), environment='latent-static-v1')
     users, items, ratings, P, R = ground_truth_matrix_to_dataset(
         truth, quantization='onetofive', bias='popularity')
 
-    print(correlation(P, truth, correlation='spearman'))
+    print(correlation(P, truth, correlation='kerdalltau'))

### 2022-06-18 data_utils.py
Try setting random seed

@@ -1,5 +1,5 @@
 import numpy as np
-from scipy.stats import pearsonr, spearmanr, kendalltau
+from scipy.stats import pearsonr, spearmanr
 from random import random
 from math import exp
 import reclab
@@ -54,6 +54,7 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
         P: P[i, j] is the probability of observing rating i on item j.
         R: R is a quantized version of the input matrix. Noise is also added.
     """
+
     m, n = matrix.shape
     np.random.shuffle(matrix)
 
@@ -172,14 +173,12 @@ def correlation(P, matrix, correlation='pearson'):
         return pearsonr(P.flatten(), matrix.flatten())[0]
     elif correlation == 'spearman':
         return spearmanr(P.flatten(), matrix.flatten())[0]
-    elif correlation == 'kendalltau':
-        return kendalltau(P.flatten(), matrix.flatten())[0]
 
 
 if __name__ == '__main__':
     truth = generate_ground_truth_matrix(
-        (1000, 1000), environment='latent-static-v1')
+        (1000, 1000), environment='ml-100k-v1')
     users, items, ratings, P, R = ground_truth_matrix_to_dataset(
         truth, quantization='onetofive', bias='popularity')
 
-    print(correlation(P, truth, correlation='kerdalltau'))
+    print(correlation(P, truth, correlation='spearman'))

### 2022-06-20 data_utils.py
Convert observed dataset to surprise format

@@ -1,4 +1,5 @@
 import numpy as np
+import pandas as pd
 from scipy.stats import pearsonr, spearmanr
 from random import random
 from math import exp
@@ -174,6 +175,19 @@ def correlation(P, matrix, correlation='pearson'):
     elif correlation == 'spearman':
         return spearmanr(P.flatten(), matrix.flatten())[0]
 
+
+def to_dataframe(ratings):
+    ratings_dict = {
+        'itemID': [],
+        'userID': [],
+        'rating': []
+    }
+    for key, value in ratings.items():
+        userID, itemID = key
+        ratings_dict['itemID'].append(itemID)
+        ratings_dict['userID'].append(userID)
+        ratings_dict['rating'].append(value)
+    return pd.DataFrame(ratings_dict)
 
 if __name__ == '__main__':
     truth = generate_ground_truth_matrix(

### 2022-06-20 paper/paper.tex
Try setting random seed
Explain bias loses information

@@
 Assume the (flattened) ground truth ratings is a standard Gaussian with differential entropy
-\begin{align*} \HH(R\sim N(0, 1)) & = -\EE[\log R]                    \\ & = -\int_{-\infty}^\infty \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}} \log \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}} \, dx \\
+\begin{align*} \HH(R\sim N(0, 1)) & = -\int_{-\infty}^\infty \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}} \log \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}} \, dx \\
                                   & = \frac{1}{2}\log 2\pi e                                                                                            \\
@@
 \[ \HH(R) > \HH(Q_2) > \HH(Q_1) \]
+Next, we can prove that selection bias causes information loss.
+Assuming the selection bias causes ratings to have unequal probability to be sampled, and higher ratings are more likely to be sampled.
+The distribution of the observed ratings would then be right-skewed.
+We prove that, in the binary case, this reduces the information loss.
+Let $p$ be the probability that the sampled rating is positive.
+\[\HH = p\log p + (1-p) \log (1-p) \]
+\[ \frac{d\HH}{dp} = 1 + \log p + \frac{1-p}{p-1} - \log(1-p) = 0 \]
+which is optimized when $p = 0.5$.
+This result extends to all $n$-quantizations (can be proved using Lagrange multipliers.) See \url{http://pillowlab.princeton.edu/teaching/statneuro2018/slides/notes08_infotheory.pdf}.
 
-Next: discussion
 \end{document}

### 2022-06-20 data_utils.py

@@ -182,17 +182,18 @@ def to_dataframe(ratings):
         'userID': [],
         'rating': []
     }
-    for key, value in ratings.items():
+    for key, value in ratings.itsems():
         userID, itemID = key
         ratings_dict['itemID'].append(itemID)
         ratings_dict['userID'].append(userID)
         ratings_dict['rating'].append(value)
     return pd.DataFrame(ratings_dict)
 
+
 if __name__ == '__main__':
     truth = generate_ground_truth_matrix(
         (1000, 1000), environment='ml-100k-v1')
     users, items, ratings, P, R = ground_truth_matrix_to_dataset(
         truth, quantization='onetofive', bias='popularity')
-
-    print(correlation(P, truth, correlation='spearman'))
+    df = to_dataframe(ratings)
+    print(df.head())

### 2022-06-21 data_utils.py
Import surprise

@@ -182,7 +182,7 @@ def to_dataframe(ratings):
         'userID': [],
         'rating': []
     }
-    for key, value in ratings.itsems():
+    for key, value in ratings.items():
         userID, itemID = key
         ratings_dict['itemID'].append(itemID)
         ratings_dict['userID'].append(userID)

### 2022-06-26 data_utils.py
Be able to calculate RMSE

@@ -166,14 +166,14 @@ def sample(value, sample_prob=0.1):
         return value
 
 
-def correlation(P, matrix, correlation='pearson'):
-    """
-    Helper function that computes the correlation between two matrices.
-    """
-    if correlation == 'pearson':
-        return pearsonr(P.flatten(), matrix.flatten())[0]
-    elif correlation == 'spearman':
-        return spearmanr(P.flatten(), matrix.flatten())[0]
+# def correlation(P, matrix, correlation='pearson'):
+#     """
+#     Helper function that computes the correlation between two matrices.
+#     """
+#     if correlation == 'pearson':
+#         return pearsonr(P.flatten(), matrix.flatten())[0]
+#     elif correlation == 'spearman':
+#         return spearmanr(P.flatten(), matrix.flatten())[0]
 
 
 def to_dataframe(ratings):
@@ -183,11 +183,26 @@ def to_dataframe(ratings):
         'rating': []
     }
     for key, value in ratings.items():
-        userID, itemID = key
-        ratings_dict['itemID'].append(itemID)
-        ratings_dict['userID'].append(userID)
-        ratings_dict['rating'].append(value)
+        if value is not None:
+            userID, itemID = key
+            ratings_dict['itemID'].append(itemID)
+            ratings_dict['userID'].append(userID)
+            ratings_dict['rating'].append(value)
     return pd.DataFrame(ratings_dict)
+
+
+def generate_test_dataframe(R):
+    test_dict = {'itemID': [],
+        'userID': [],
+        'rating': []}
+    m, n = R.shape
+    for i in range(m):
+        for j in range(n):
+            if random() > 0.99:
+                test_dict['itemID'].append(j)
+                test_dict['userID'].append(i)
+                test_dict['rating'].append(R[i][j])
+    return pd.DataFrame(test_dict)
 
 
 if __name__ == '__main__':

### 2022-06-28 data_utils.py
Preliminary results

@@ -62,19 +62,54 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
     R = matrix.copy()
 
     # Normalize the ground truth and add Gaussian noise
-    R = (R - R.min()) / (R.max() - R.min())
-    R = R + np.random.normal(0, noise, R.shape)
+    R_no_noise = (R - R.min()) / (R.max() - R.min())
+    R = R_no_noise + np.random.normal(0, noise, R.shape)
 
     # Quantize the matrix
     if quantization == 'binary':
         R[R <= 0.5] = 0
+        R_no_noise[R_no_noise <= 0.5] = 0
         R[R > 0.5] = 1
+        R_no_noise[R_no_noise > 0.5] = 1
     elif quantization == 'onetofive':
         R[R <= 0.2] = 1
+        R_no_noise[R_no_noise <= 0.2] = 1
         R[R <= 0.4] = 2
+        R_no_noise[R_no_noise <= 0.4] = 2
         R[R <= 0.6] = 3
+        R_no_noise[R_no_noise <= 0.6] = 3
         R[R <= 0.8] = 4
+        R_no_noise[R_no_noise <= 0.8] = 4
         R[R <= 1] = 5
+        R_no_noise[R_no_noise <= 1] = 5
+    elif quantization == 'onetothree':
+        R[R <= 0.33] = 1
+        R_no_noise[R_no_noise <= 0.33] = 1
+        R[R <= 0.66] = 2
+        R_no_noise[R_no_noise <= 0.66] = 2
+        R[R <= 1] = 3
+        R_no_noise[R_no_noise <= 1] = 3  
+    elif quantization == 'onetoten':
+        R[R <= 0.1] = 1
+        R_no_noise[R_no_noise <= 0.1] = 1
+        R[R <= 0.2] = 2
+        R_no_noise[R_no_noise <= 0.2] = 2
+        R[R <= 0.3] = 3
+        R_no_noise[R_no_noise <= 0.3] = 3
+        R[R <= 0.4] = 4
+        R_no_noise[R_no_noise <= 0.4] = 4
+        R[R <= 0.5] = 5
+        R_no_noise[R_no_noise <= 0.5] = 5
+        R[R <= 0.6] = 6
+        R_no_noise[R_no_noise <= 0.6] = 6
+        R[R <= 0.7] = 7
[…]

### 2022-06-30 paper/neurips_2022.tex
Draft an introduction

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2022

% ready for submission
\usepackage{neurips_2022}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2022}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2022}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2022}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

\title{Formatting Instructions For NeurIPS 2022}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  David S.~Hippocampus\thanks{Use footnote for providing further information
    about author (webpage, alternative address)---\emph{not} for acknowledging
    funding agencies.} \\
  Department of Computer Science\\
  Cranberry-Lemon University\\
  Pittsburgh, PA 15213 \\
  \texttt{hippo@cs.cranberry-lemon.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
  both the left- and right-hand margins. Use 10~point type, with a vertical
  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
  bold, and in point size 12. Two line spaces precede the abstract. The abstract
  must be limited to one paragraph.
\end{abstract}

\section{}

Please read the instructions below carefully and follow them faithfully.

\subsection{Style}

Papers to be submitted to NeurIPS 2022 must be prepared according to the
instructions presented here. Papers may only be up to {\bf nine} pages long,
including figures. Additional pages \emph{containing only acknowledgments and
references} are allowed. Papers that exceed the page limit will not be
reviewed, or in any other way considered for presentation at the conference.

The margins in 2022 are the same as those in 2007, which allow for $\sim$$15\%$
more words in the paper compared to earlier years.

Authors are required to use the NeurIPS \LaTeX{} style files obtainable at the
NeurIPS website as indicated below. Please make sure you use the current files
and not previous versions. Tweaking the style files may be grounds for
rejection.

\subsection{Retrieval of style files}

The style files for NeurIPS and other conference information are available on
the World Wide Web at
\begin{center}
  \url{http://www.neurips.cc/}
\end{center}
The file \verb+neurips_2022.pdf+ contains these instructions and illustrates the
various formatting requirements your NeurIPS paper must satisfy.

The only supported style file for NeurIPS 2022 is \verb+neurips_2022.sty+,
rewritten for \LaTeXe{}.  \textbf{Previous style files for \LaTeX{} 2.09,
  Microsoft Word, and RTF are no longer supported!}

The \LaTeX{} style file contains three optional arguments: \verb+final+, which
creates a camera-ready copy, \verb+preprint+, which creates a preprint for
submission to, e.g., arXiv, and \verb+nonatbib+, which will not load the
\verb+natbib+ package for you in case of package clash.

\paragraph{Preprint option}
If you wish to post a preprint of your work online, e.g., on arXiv, using the
NeurIPS style, please use the \verb+preprint+ option. This will create a
nonanonymized version of your work with the text ``Preprint. Work in progress.''
in the footer. This version may be distributed as you see fit. Please \textbf{do
  not} use the \verb+final+ option, which should \textbf{only} be used for
papers accepted to NeurIPS.

At submission time, please omit the \verb+final+ and \verb+preprint+
options. This will anonymize your submission and add line numbers to aid
review. Please do \emph{not} refer to these line numbers in your paper as they
will be removed during generation of camera-ready copies.

The file \verb+neurips_2022.tex+ may be used as a ``shell'' for writing your
paper. All you have to do is replace the author, title, abstract, and text of
the paper with your own.

The formatting instructions contained in these style files are summarized in
Sections \ref{gen_inst}, \ref{headings}, and \ref{others} below.

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas) wide and
9~inches (54~picas) long. The left margin is 1.5~inch (9~picas).  Use 10~point
type with a vertical spacing (leading) of 11~points.  Times New Roman is the
preferred typeface throughout, and will be selected for you by default.
Paragraphs are separated by \nicefrac{1}{2}~line space (5.5 points), with no
indentation.

The paper title should be 17~point, initial caps/lower case, bold, centered
between two horizontal rules. The top rule should be 4~points thick and the
bottom rule should be 1~point thick. Allow \nicefrac{1}{4}~inch space above and
below the title to rules. All pages should start at 1~inch (6~picas) from the
top of the page.

For the final version, authors' names are set in boldface, and each name is
centered above the corresponding address. The lead author's name is to be listed
first (left-most), and the co-authors' names (if different address) are set to
follow. If there is only one co-author, list both author and co-author side by
side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

All headings should be lower case (except for first word and proper nouns),
flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the heading in
bold, flush left, and inline with the text, with the heading followed by 1\,em
of space.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.  Citations may be
author/year or numeric, as long as you maintain internal consistency.  As to the
format of the references themselves, any style is acceptable as long as it is
used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations appropriate for
use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may add the
following before loading the \verb+neurips_2022+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add the optional
argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{neurips_2022}
\end{verbatim}

As submission is double blind, refer to your own published work in the third
person. That is, use ``In the previous work of Jones et al.\ [4],'' not ``In our
previous work [4].'' If you cite your other papers that are not widely available
(e.g., a journal paper under review), use anonymous author names in the
citation, e.g., an author of the form ``A.\ Anonymous.''

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote, indicate
footnotes with a number\footnote{Sample of the first footnote.} in the
text. Place the footnotes at the bottom of the page on which they appear.
Precede the footnote with a horizontal rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

All artwork must be neat, clean, and legible. Lines should be dark enough for
purposes of reproduction. The figure number and caption always appear after the
figure. Place one line space before the figure caption and one line space after
the figure. The figure caption should be lower case (except for first word and
proper nouns); figures are numbered consecutively.

You may use color figures.  However, it is best for the figure captions and the
paper body to be legible if the paper is printed in either black/white or in
color.

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table number and
title always appear before the table.  See Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical rules.} We
strongly suggest the use of the \verb+booktabs+ package, which allows for
typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Final instructions}

Do not change any aspects of the formatting parameters in the style files.  In
particular, do not modify the width or length of the rectangle the text should
fit into, and do not change font sizes (except perhaps in the
\textbf{References} section; see below). Please note that pages should be
numbered.

\section{Preparing PDF files}

Please prepare submission files with paper size ``US Letter,'' and not, for
example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file must only
contain Type 1 or Embedded TrueType fonts. Here are a few instructions to
achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader, select the
  menu Files$>$Document Properties$>$Fonts and select Show All Fonts. You can
  also use the program \verb+pdffonts+ which comes with \verb+xpdf+ and is
  available out-of-the-box on most Linux machines.

\item The IEEE has recommendations for generating PDF files whose fonts are also
  acceptable for NeurIPS. Please see
  \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item \verb+xfig+ "patterned" shapes are implemented with bitmap fonts.  Use
  "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You should use
  the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or \verb+\mathbb{C}+
for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You can also use the following
workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the \verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we will ask
you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always specify the
figure width as a multiple of the line width as in the example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly hyphenate a
line. Please give LaTeX hyphenation hints using the \verb+\-+ command when
necessary.

\begin{ack}
Use unnumbered first level headings for the acknowledgments. All acknowledgments
go at the end of the paper before the list of references. Moreover, you are required to declare
funding (financial activities supporting the submitted work) and competing interests (related financial activities outside the submitted work).
More information about this disclosure can be found at: \url{https://neurips.cc/Conferences/2022/PaperInformation/FundingDisclosure}.

Do {\bf not} include this section in the anonymized submission, only in the final paper. You can use the \texttt{ack} environment provided in the style file to autmoatically hide this section in the anonymized submission.
\end{ack}

\section*{References}

References follow the acknowledgments. Use unnumbered first-level heading for
the references. Any choice of citation style is acceptable as long as you are
consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
when listing the references.
Note that the Reference section does not count towards the page limit.
\medskip

{
\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
(eds.), {\it Advances in Neural Information Processing Systems 7},
pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
  Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
recall at excitatory recurrent synapses and cholinergic modulation in rat
hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Checklist}

%%% BEGIN INSTRUCTIONS %%%
The checklist follows the references.  Please
read the checklist guidelines carefully for information on how to answer these
questions.  For each question, change the default \answerTODO{} to \answerYes{},
\answerNo{}, or \answerNA{}.  You are strongly encouraged to include a {\bf
justification to your answer}, either by referencing the appropriate section of
your paper or providing a brief inline description.  For example:
\begin{itemize}
  \item Did you include the license to the code and datasets? \answerYes{See Section~\ref{gen_inst}.}
  \item Did you include the license to the code and datasets? \answerNo{The code and the data are proprietary.}
  \item Did you include the license to the code and datasets? \answerNA{}
\end{itemize}
Please do not modify the questions and only use the provided macros for your
answers.  Note that the Checklist section does not count towards the page
limit.  In your paper, please delete this instructions block and only keep the
Checklist section heading above along with the questions/answers below.
%%% END INSTRUCTIONS %%%

\begin{enumerate}

\item For all authors...
\begin{enumerate}
  \item Do the main claims made in the abstract and introduction accurately reflect the paper's contributions and scope?
    \answerTODO{}
  \item Did you describe the limitations of your work?
    \answerTODO{}
  \item Did you discuss any potential negative societal impacts of your work?
    \answerTODO{}
  \item Have you read the ethics review guidelines and ensured that your paper conforms to them?
    \answerTODO{}
\end{enumerate}

\item If you are including theoretical results...
\begin{enumerate}
  \item Did you state the full set of assumptions of all theoretical results?
    \answerTODO{}
        \item Did you include complete proofs of all theoretical results?
    \answerTODO{}
\end{enumerate}

\item If you ran experiments...
\begin{enumerate}
  \item Did you include the code, data, and instructions needed to reproduce the main experimental results (either in the supplemental material or as a URL)?
    \answerTODO{}
  \item Did you specify all the training details (e.g., data splits, hyperparameters, how they were chosen)?
    \answerTODO{}
        \item Did you report error bars (e.g., with respect to the random seed after running experiments multiple times)?
    \answerTODO{}
        \item Did you include the total amount of compute and the type of resources used (e.g., type of GPUs, internal cluster, or cloud provider)?
    \answerTODO{}
\end{enumerate}

\item If you are using existing assets (e.g., code, data, models) or curating/releasing new assets...
\begin{enumerate}
  \item If your work uses existing assets, did you cite the creators?
    \answerTODO{}
  \item Did you mention the license of the assets?
    \answerTODO{}
  \item Did you include any new assets either in the supplemental material or as a URL?
    \answerTODO{}
  \item Did you discuss whether and how consent was obtained from people whose data you're using/curating?
    \answerTODO{}
  \item Did you discuss whether the data you are using/curating contains personally identifiable information or offensive content?
    \answerTODO{}
\end{enumerate}

\item If you used crowdsourcing or conducted research with human subjects...
\begin{enumerate}
  \item Did you include the full text of instructions given to participants and screenshots, if applicable?
    \answerTODO{}
  \item Did you describe any potential participant risks, with links to Institutional Review Board (IRB) approvals, if applicable?
    \answerTODO{}
  \item Did you include the estimated hourly wage paid to participants and the total amount spent on participant compensation?
    \answerTODO{}
\end{enumerate}

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Appendix}

Optionally include extra information (complete proofs, additional experiments and plots) in the appendix.
This section will often be part of the supplemental material.

\end{document}

### 2022-06-30 paper/paper.tex
Run experiment and plot graph
Preliminary results
Draft an introduction

@@
-% My template
-\documentclass[12pt, letterpaper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[amsthm]{ntheorem}
+\documentclass{article}
+
+% if you need to pass options to natbib, use, e.g.:
+%     \PassOptionsToPackage{numbers, compress}{natbib}
+% before loading neurips_2022
+
+% ready for submission
+\usepackage{neurips_2022}
+
+% to compile a preprint version, e.g., for submission to arXiv, add add the
+% [preprint] option:
+%     \usepackage[preprint]{neurips_2022}
+
+% to compile a camera-ready version, add the [final] option, e.g.:
+%     \usepackage[final]{neurips_2022}
+
+% to avoid loading the natbib package, add option nonatbib:
+%    \usepackage[nonatbib]{neurips_2022}
+
+\usepackage[utf8]{inputenc} % allow utf-8 input
+\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
+\usepackage{hyperref}       % hyperlinks
+\usepackage{url}            % simple URL typesetting
+\usepackage{booktabs}       % professional-quality tables
+\usepackage{amsfonts}       % blackboard math symbols
+\usepackage{nicefrac}       % compact symbols for 1/2, etc.
+\usepackage{microtype}      % microtypography
+\usepackage{xcolor}         % colors
 \usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{url}
-\usepackage{enumitem}
-\usepackage{hyperref}
-\hypersetup{
-    colorlinks,
-    citecolor=black,
-    filecolor=black,
-    linkcolor=black,
-    urlcolor=black
-}
-\widowpenalty10000
-\clubpenalty10000
-\setlength{\parskip}{\baselineskip}
-\setlength{\parindent}{0pt}
-\setlist[enumerate]{topsep=0pt}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-\theoremstyle{definition}\newtheorem*{definition}{Definition}
-\theoremstyle{definition}\newtheorem*{example}{Example}
-\theoremstyle{definition}\newtheorem*{remark}{Remark}
+
 \newcommand{\vect}[1]{\boldsymbol{#1}}
@@
 \def\Mm{\mathcal{M}}
-\DeclareMathOperator*{\argmax}{arg\,max}
-\DeclareMathOperator*{\argmin}{arg\,min}
 
-\title{\vspace{-2.0cm}Mitigating the Effect of Sampling Bias on Recommender Systems through Quantization}
-\author{Fengyu Li, Sarah Dean}
-\date{}
+\title{Title}
+
+% The \author macro works with any number of authors. There are two commands
+% used to separate the names and addresses of multiple authors: \And and \AND.
+%
+% Using \And between authors leaves it to LaTeX to determine where to break the
+% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
+% authors names on the first line, and the last on the second line, try using
+% \AND instead of \And before the third author name.
+
+\author{%
+  David S.~Hippocampus\thanks{Use footnote for providing further information
+    about author (webpage, alternative address)---\emph{not} for acknowledging
+    funding agencies.} \\
+  Department of Computer Science\\
+  Cranberry-Lemon University\\
+  Pittsburgh, PA 15213 \\
+  \texttt{hippo@cs.cranberry-lemon.edu} \\
+  % examples of more authors
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+  % \AND
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+  % \And
+  % Coauthor \\
+  % Affiliation \\
+  % Address \\
+  % \texttt{email} \\
+}
 
 \begin{document}
+
 \maketitle
@@
 \begin{abstract}
-    Placleholder.
+  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
+  both the left- and right-hand margins. Use 10~point type, with a vertical
+  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
+  bold, and in point size 12. Two line spaces precede the abstract. The abstract
+  must be limited to one paragraph.
 \end{abstract}
@@
 \section{Introduction}
+Selection bias is one of the most prevalent source of biases in a dataset for recommender systems.
+Selection bias happens when there is a pattern in the items that a user rates, either implicitly or explicitly.
+For example, in a recommender system for movies, users might only watch and rate the movies that are attractive to them in the first place.
+Selection bias is particularly relevant to recommender systems because when the system actively recommends items that the users like, it is creating bias by preventing users from accessing uncovered items.
+Selection bias introduces a Missing Not At Random (MNAR) condition to the observed dataset by conditioning it on the predictions of the recommender system.
+Tackling the selection bias in a recommender dataset has been a constant challenge in designing recommender algorithms
+
+Quantization is a practical way to categorize an input from a continuous distribution to a discrete set of bins.
+We assume the ground truth of a user's rating is a continuous distribution over items because of the inherent impossibility of numerically expressing the feeling of enjoyment and excitement.
+We treat a recommender dataset as a noised quantization of the ground truth.
+In other words, users convert their feelings into one of the discrete choices when they rate an item.
+Recommender datasets are quantized into different degrees.
+For example, a typical movie rating dataset uses a 5-quantization.
+A tinder like/dislike dataset uses a 2-quantization.
+
+We hypothesize that quantization and the effect of selection bias on the recommender system is positively correlated: the more quantized the dataset is (the fewer bins there are), the more severe the selection bias becomes in terms of the recommender system's performance.
+Therefore, a good-performing recommender system should prioritize training data that is less quantized.
+
+In reality, however, less quantized data might only represent a small subset of all the data one is capable of collecting.
+This is because less quantized data is typically explicit while more quantized data is typically implicit.
+That is to say, less quantized data would require a user explicitly rating an item, while more quantized data would be more easily gathered by observing the users' behaviors.
+For example, by observing the movie a user spontaneously chooses to see or if a user watches a movie's trailer to the end, we can easily create a 1-quantized or 2-quantized dataset, correspondingly, without the user paying any effort.
+In this paper, we propose a novel framework that utilizes these data.
+Our framework specifically exploits the fact that more quantized data is more susceptible to selection bias and treats it as an advantage that helps less quantized dataset better overcome the damage caused by selection bias.
+
+The main contribution of this work is three-fold:
+\begin{enumerate}
+  \item We design experiments under a simulated environment and shows that a more quantized dataset is more susceptible to selection bias (and thus the MNAR condition.)
+  \item We approach the above empirical results from a theoretical lens and explains the complex relationship between quantization, selection bias, and a recommender's performance using tools from information theory.
+  \item We propose a cross-dataset learning framework that empowers existing recommender algorithms to make use of the more quantized, implicit data. \textcolor{red}{EXPLAIN MORE HERE}
+\end{enumerate}
+
+Our hypothesis and theoretical results are verified through a series of empirical evaluation in a simulated environment where users' behaviors are carefully modeled.
+We show that our learning framework is significantly better than baseline recommender systems that neglects selection bias.
+We also show that our methods are substantially better than its single-dataset counterpart.
+\textcolor{red}{EXPLAIN MORE HERE}
+
+\section{Related Work}
+
+\subsection{Recommender Systems Against Selection Bias}
+
+\subsection{Quantization}
+
+\section{Theoretical Analysis}
+
+\section{Cross-Dataset Learning Framework}
+
+\section{Experiment}
+We designed experiments under simulated environments that verify the results introduced above.
+We first verify that more quantized dataset is more susceptible to selection bias and MNAR.
+We then runs our learning framework and compares it with multiple baselines.
+
+\subsection{Experiment Setup}
+
+\subsection{Controlling Bias}
+
+\subsection{Results}
+
+\section{Conclusion}
 
@@
 Assume the (flattened) ground truth ratings is a standard Gaussian with differential entropy
-\begin{align*} \HH(R\sim N(0, 1)) & = -\int_{-\infty}^\infty \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}} \log \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}} \, dx \\
+\begin{align} \HH(R\sim N(0, 1))
    & = \frac{1}{2}\log 2\pi e \\
    & = 2.05
-\end{align*}
+\end{align}
 We can show that quantization (discretization) loses information. Consider
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \end{document}

### 2022-06-30 paper/paper_plain.tex
Draft an introduction

--- 2022-06-30 paper/paper.tex
@@
-\documentclass{article}
-
-% if you need to pass options to natbib, use, e.g.:
-%     \PassOptionsToPackage{numbers, compress}{natbib}
-% before loading neurips_2022
-
-% ready for submission
-\usepackage{neurips_2022}
-
-% to compile a preprint version, e.g., for submission to arXiv, add add the
-% [preprint] option:
-%     \usepackage[preprint]{neurips_2022}
-
-% to compile a camera-ready version, add the [final] option, e.g.:
-%     \usepackage[final]{neurips_2022}
-
-% to avoid loading the natbib package, add option nonatbib:
-%    \usepackage[nonatbib]{neurips_2022}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{xcolor}         % colors
+% My template
+\documentclass[12pt, letterpaper]{article}
+\usepackage[utf8]{inputenc}
+\usepackage[amsthm]{ntheorem}
 \usepackage{amsmath}
-
+\usepackage{amsfonts}
+\usepackage{url}
+\usepackage{parskip}
+\usepackage{enumitem}
+\usepackage{hyperref}
+\hypersetup{
+    colorlinks,
+    citecolor=black,
+    filecolor=black,
+    linkcolor=black,
+    urlcolor=black
+}
+\setlist[enumerate]{topsep=0pt}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[theorem]
+\newtheorem{lemma}[theorem]{Lemma}
+\theoremstyle{definition}\newtheorem*{definition}{Definition}
+\theoremstyle{definition}\newtheorem*{example}{Example}
+\theoremstyle{definition}\newtheorem*{remark}{Remark}
 \newcommand{\vect}[1]{\boldsymbol{#1}}
@@
 \def\Mm{\mathcal{M}}
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
-\title{Title}
-
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to LaTeX to determine where to break the
-% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
-% authors names on the first line, and the last on the second line, try using
-% \AND instead of \And before the third author name.
-
-\author{%
-  David S.~Hippocampus\thanks{Use footnote for providing further information
-    about author (webpage, alternative address)---\emph{not} for acknowledging
-    funding agencies.} \\
-  Department of Computer Science\\
-  Cranberry-Lemon University\\
-  Pittsburgh, PA 15213 \\
-  \texttt{hippo@cs.cranberry-lemon.edu} \\
-  % examples of more authors
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \AND
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-}
+\title{\vspace{-2.0cm}Title }
+\author{Fengyu Li, Sarah Dean}
+\date{}
 
 \begin{document}
-
 \maketitle
@@
 \begin{abstract}
-  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
-  both the left- and right-hand margins. Use 10~point type, with a vertical
-  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
-  bold, and in point size 12. Two line spaces precede the abstract. The abstract
-  must be limited to one paragraph.
+
 \end{abstract}
@@
 Therefore, a good-performing recommender system should prioritize training data that is less quantized.
-
 In reality, however, less quantized data might only represent a small subset of all the data one is capable of collecting.
@@
     \item We approach the above empirical results from a theoretical lens and explains the complex relationship between quantization, selection bias, and a recommender's performance using tools from information theory.
-  \item We propose a cross-dataset learning framework that empowers existing recommender algorithms to make use of the more quantized, implicit data. \textcolor{red}{EXPLAIN MORE HERE}
+    \item We propose a cross-dataset learning framework that empowers existing recommender algorithms to make use of the more quantized, implicit data. EXPLAIN MORE HERE
 \end{enumerate}
@@
 We also show that our methods are substantially better than its single-dataset counterpart.
-\textcolor{red}{EXPLAIN MORE HERE}
+EXPLAIN MORE HERE
 
@@
 Assume the (flattened) ground truth ratings is a standard Gaussian with differential entropy
-\begin{align} \HH(R\sim N(0, 1))
+\begin{align*} \HH(R\sim N(0, 1)) & = -\int_{-\infty}^\infty \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}} \log \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}} \, dx \\
                                   & = \frac{1}{2}\log 2\pi e                                                                                            \\
                                   & = 2.05
-\end{align}
+\end{align*}
 We can show that quantization (discretization) loses information. Consider
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \end{document}

### 2022-07-21 paper/paper.tex
Add algo.py

@@
 
-\title{Title}
+\title{Propensity Estimation Matrix Factorization Through Mixed-Quantized Dataset}
 
@@
 
-The main contribution of this work is three-fold:
+The main contribution of this work is two-fold:
 \begin{enumerate}
   \item We design experiments under a simulated environment and shows that a more quantized dataset is more susceptible to selection bias (and thus the MNAR condition.)
-  \item We approach the above empirical results from a theoretical lens and explains the complex relationship between quantization, selection bias, and a recommender's performance using tools from information theory.
   \item We propose a cross-dataset learning framework that empowers existing recommender algorithms to make use of the more quantized, implicit data. \textcolor{red}{EXPLAIN MORE HERE}
@@
 \subsection{Quantization}
-
-\section{Theoretical Analysis}
 

### 2022-07-21 algo.py

@@ -1,0 +1,263 @@
+from __future__ import (absolute_import, division, print_function,
+                        unicode_literals)
+from data_utils import generate_ground_truth_matrix, ground_truth_matrix_to_dataset, to_dataframe, generate_test_dataframe
+from surprise import Reader
+from surprise import Dataset
+from surprise.model_selection import train_test_split
+from surprise.model_selection import cross_validate
+from surprise import accuracy
+import numpy as np
+from surprise.prediction_algorithms import AlgoBase
+from surprise.prediction_algorithms import PredictionImpossible
+from surprise.utils import get_rng
+
+
+class SVD(AlgoBase):
+
+    def __init__(self, n_factors=100, n_epochs=20, biased=True, init_mean=0,
+                 init_std_dev=.1, lr_all=.005,
+                 reg_all=.02, lr_bu=None, lr_bi=None, lr_pu=None, lr_qi=None,
+                 reg_bu=None, reg_bi=None, reg_pu=None, reg_qi=None,
+                 random_state=None, verbose=False):
+
+        self.n_factors = n_factors
+        self.n_epochs = n_epochs
+        self.biased = biased
+        self.init_mean = init_mean
+        self.init_std_dev = init_std_dev
+        self.lr_bu = lr_bu if lr_bu is not None else lr_all
+        self.lr_bi = lr_bi if lr_bi is not None else lr_all
+        self.lr_pu = lr_pu if lr_pu is not None else lr_all
+        self.lr_qi = lr_qi if lr_qi is not None else lr_all
+        self.reg_bu = reg_bu if reg_bu is not None else reg_all
+        self.reg_bi = reg_bi if reg_bi is not None else reg_all
+        self.reg_pu = reg_pu if reg_pu is not None else reg_all
+        self.reg_qi = reg_qi if reg_qi is not None else reg_all
+        self.random_state = random_state
+        self.verbose = verbose
+
+        AlgoBase.__init__(self)
+
+    def fit(self, trainset):
+
+        AlgoBase.fit(self, trainset)
+        self.sgd(trainset)
+
+        return self
+
+    def sgd(self, trainset):
+
+        u, i, f = 0, 0, 0
+        r, err, dot, puf, qif = 0., 0., 0., 0., 0.
+        global_mean = self.trainset.global_mean
+
+        lr_bu = self.lr_bu
+        lr_bi = self.lr_bi
+        lr_pu = self.lr_pu
[…]

### 2022-07-21 data_utils.py
Fix bugs in masked propensity estimation

@@ -7,21 +7,8 @@ import reclab
 
 EPSILON = 1e-8
 
-
 def generate_ground_truth_matrix(dimensions, environment='random'):
-    """
-    Generates a ground truth matrix.
-
-    Parameters:
-        environment: the behavoral model that establishes the ground truth matrix. A list of predefined, supported
-        environments is available at https://github.com/berkeley-reclab/RecLab/blob/master/reclab/environments/registry.py.
-        dimensions: a tuple containing the dimensions of the matrix.
-
-    Returns:
-        a mxn matrix containing the (continuously distributed) ground truth for each pair of m users and n items.
-    """
     m, n = dimensions
-
     if environment == 'random':
         return np.random.rand(dimensions)
     elif environment == 'ml-100k-v1':
@@ -35,30 +22,8 @@ def generate_ground_truth_matrix(dimensions, environment='random'):
 
 
 def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=None, beta=1, noise=0.05):
-    """
-    Converts a ground truth matrix to a recommender dataset. First simulate the observations on the ground truth
-    matrix, then convert them to a recommender dataset.
-    Bias can be introduced in the dataset.
-
-    Parameters:
-        matrix: a mxn matrix that contains the ground truth for each pair of m users and n items.
-        quantization: the way to quantize / discretize the ratings.
-        sample_prob: the expected probability of sampling a chosen pair of users and items.
-        bias: the type of sampling bias.
-        beta: hyperparameter that controls the degress of the bias.
-        noise: hyperparameter that controls the degree of noise.
-
-    Returns:
-        users: a dictionary whose ith entry is a list of observed ratings for the ith user.
-        itmes: a dictionary whose ith entry is a list of observed ratings for the ith item.
-        ratings: a dictionary whose entry with key (i, j) is the observed rating for user i on item j.
-        P: P[i, j] is the probability of observing rating i on item j.
-        R: R is a quantized version of the input matrix. Noise is also added.
-    """
-
     m, n = matrix.shape
     np.random.shuffle(matrix)
-
     R = matrix.copy()
 
     # Normalize the ground truth and add Gaussian noise
@@ -197,16 +162,6 @@ def sample(value, sample_prob=0.1):
         return value
 
 
-# def correlation(P, matrix, correlation='pearson'):
-#     """
[…]

### 2022-07-29 paper/paper.tex
Add algo

@@
 \def\Mm{\mathcal{M}}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 
+\[ \argmin_{V,W,A} \frac{1}{N}\left(\frac{(Y_{u,i}- (V_u^TW_i + A))^2}{P_{u,i}} \right) + c \Vert A \Vert^2\]
+where $A = \{ b_u, b_i, \mu \}$ represents the standard bias parameters (offset), $\hat{Y} = V_u^TW_i + A$ is the predicted rating, and $c \Vert A \Vert^2$ is the regularizer.
+
+Stochastic gradient descent (SGD) step for one sample rating $(i,j)$:
+\[ b_u \gets b_u - \alpha(b_u - \frac{1}{P_{i,j}}(r_{i,j} - \hat{r}_{i,j})) \]
+\[ b_i \gets b_i - \alpha(b_i - \frac{1}{P_{i,j}}(r_{i,j} - \hat{r}_{i,j})) \]
+\[ V_i \gets V_i - \alpha(V_i - \frac{1}{P_{i,j}}W_j(r_{i,j} - \hat{r}_{i,j})) \]
+\[ W_j \gets W_j - \alpha(W_j - \frac{1}{P_{i,j}}V_i(r_{i,j} - \hat{r}_{i,j})) \]
+where $\alpha$ is the learning rate, which can be tuned independently for each parameter. In our experiments, we fix $\alpha$ across parameters.
+
 \section{Experiment}

### 2022-07-29 algo.py

@@ -1,16 +1,17 @@
 from __future__ import (absolute_import, division, print_function,
                         unicode_literals)
-from data_utils import generate_ground_truth_matrix, ground_truth_matrix_to_dataset, to_dataframe, generate_test_dataframe
+
+# import os
+# os.environ['KMP_DUPLICATE_LIB_OK']='True'
+                        
+from data_utils import generate_ground_truth_matrix, ground_truth_matrix_to_dataset, to_dataframe, generate_test_dataframe, masked_nb_propensity_estimation
+import numpy as np
 from surprise import Reader
 from surprise import Dataset
-from surprise.model_selection import train_test_split
-from surprise.model_selection import cross_validate
 from surprise import accuracy
-import numpy as np
 from surprise.prediction_algorithms import AlgoBase
 from surprise.prediction_algorithms import PredictionImpossible
 from surprise.utils import get_rng
-
 
 class SVD(AlgoBase):
 
@@ -128,12 +129,15 @@ class SVD(AlgoBase):
 
 class PropensitySVD(AlgoBase):
 
-    def __init__(self, n_factors=100, n_epochs=20, biased=True, init_mean=0,
+    def __init__(self, p, n_factors=100, n_epochs=20, biased=True, init_mean=0,
                  init_std_dev=.1, lr_all=.005,
                  reg_all=.02, lr_bu=None, lr_bi=None, lr_pu=None, lr_qi=None,
                  reg_bu=None, reg_bi=None, reg_pu=None, reg_qi=None,
-                 random_state=None, verbose=False, P=np.zeros(1)):
+                 random_state=None, verbose=False):
 
+        if p is None:
+            assert False, 'Please use standard SVD'
+        
         self.n_factors = n_factors
         self.n_epochs = n_epochs
         self.biased = biased
@@ -149,7 +153,7 @@ class PropensitySVD(AlgoBase):
         self.reg_qi = reg_qi if reg_qi is not None else reg_all
         self.random_state = random_state
         self.verbose = verbose
-        self.P = P
+        self.p = p
 
         AlgoBase.__init__(self)
 
@@ -201,16 +205,16 @@ class PropensitySVD(AlgoBase):
 
                 # update biases
                 if self.biased:
-                    bu[u] += lr_bu * (err - reg_bu * bu[u])
-                    bi[i] += lr_bi * (err - reg_bi * bi[i])
[…]

### 2022-07-29 data_utils.py

@@ -1,11 +1,18 @@
+import os
+os.environ['KMP_DUPLICATE_LIB_OK']='True'
+
 import numpy as np
 import pandas as pd
 from scipy.stats import pearsonr, spearmanr
 from random import random
 from math import exp
 import reclab
+import torch
+from torch.utils.data import DataLoader, TensorDataset
+from torch import nn
 
 EPSILON = 1e-8
+device = torch.device('cpu')
 
 def generate_ground_truth_matrix(dimensions, environment='random'):
     m, n = dimensions
@@ -205,8 +212,49 @@ def naive_propensity_estimation(ratings, shape, quantization='onetofive'):
     return np.zeros(shape) + count / size
 
 
-def mlp_propensity_estimation():
-    pass
+class MLP(nn.Module):
+    def __init__(self):
+        super(MLP, self).__init__()
+        self.flatten = nn.Flatten()
+        self.linear_relu_stack = nn.Sequential(
+            nn.Linear(2, 512),
+            nn.ReLU(),
+            nn.Linear(512, 512),
+            nn.ReLU(),
+            nn.Linear(512, 1),
+            nn.Sigmoid()
+        )
+
+    def forward(self, x):
+        x = self.flatten(x)
+        y = self.linear_relu_stack(y)
+        return y
+
+
+def mlp_propensity_estimation(ratings):
+    X = []
+    y = []
+    for (user, item), value in ratings.items():
+        X.append([user, item])
+        y.append(0 if value is None else 1)
+    X = torch.tensor(X)
+    y = torch.tensor(y)
+    dataset = TensorDataset(X, y)
+    train_dataloader = DataLoader(dataset, batch_size=16, shuffle=True)
+    model = MLP().to(device)        
+    learning_rate = 1e-3
+    epochs = 100
+    loss_fn = nn.CrossEntropyLoss()
+    optimizer = torch.optim.SGD(model.parameters(), lr=learning_rate)
+    for batch_no, (X_sample, y_sample) in enumerate(train_dataloader):
+        pred = model(X_sample)
+        loss = loss_fn(pred, y_sample)
+        optimizer.zero_grad()
+        loss.backward()
+        optimizer.step()
+        if batch_no == epochs:
+            break
+    return np.array(model.forward(X))
 
 
 def masked_nb_propensity_estimation(truth, ratings, shape, beta=0):
[…]

### 2022-08-04 algo.py
Settings

@@ -52,7 +52,7 @@ class SVD(AlgoBase):
         r, err, dot, puf, qif = 0., 0., 0., 0., 0.
         global_mean = self.trainset.global_mean
 
-        lr_bu = self.lr_bu
+        lr_bu = self.lr_bucod
         lr_bi = self.lr_bi
         lr_pu = self.lr_pu
         lr_qi = self.lr_qi

### 2022-08-04 data_utils.py
Remove garbages

@@ -1,18 +1,19 @@
-import os
-os.environ['KMP_DUPLICATE_LIB_OK']='True'
-
-import numpy as np
-import pandas as pd
-from scipy.stats import pearsonr, spearmanr
-from random import random
-from math import exp
-import reclab
-import torch
-from torch.utils.data import DataLoader, TensorDataset
 from torch import nn
+from torch.utils.data import DataLoader, TensorDataset
+import torch
+import reclab
+from math import exp
+from random import random
+from scipy.stats import pearsonr, spearmanr
+import pandas as pd
+import numpy as np
+import os
+os.environ['KMP_DUPLICATE_LIB_OK'] = 'True'
+
 
 EPSILON = 1e-8
 device = torch.device('cpu')
+
 
 def generate_ground_truth_matrix(dimensions, environment='random'):
     m, n = dimensions
@@ -206,7 +207,7 @@ def naive_propensity_estimation(ratings, shape, quantization='onetofive'):
         if value is not None:
             proportion[int(value / 0.2) - 1] += 1
             count += 1
-    proportion /= proportion.sum()        
+    proportion /= proportion.sum()
     m, n = shape
     size = m * n
     return np.zeros(shape) + count / size
@@ -241,7 +242,7 @@ def mlp_propensity_estimation(ratings):
     y = torch.tensor(y)
     dataset = TensorDataset(X, y)
     train_dataloader = DataLoader(dataset, batch_size=16, shuffle=True)
-    model = MLP().to(device)        
+    model = MLP().to(device)
     learning_rate = 1e-3
     epochs = 100
     loss_fn = nn.CrossEntropyLoss()
@@ -264,20 +265,20 @@ def masked_nb_propensity_estimation(truth, ratings, shape, beta=0):
         if value is not None:
             proportion[int(value / 0.2) - 1] += 1
             count += 1
-    proportion /= proportion.sum() 
+    proportion /= proportion.sum()
     m, n = shape
     size = m * n
     numerator = np.zeros(shape)
     for key, value in ratings.items():
         if value is not None:
             user, item = key
-            numerator[user][item] = proportion[int(value / 0.2) - 1] 
-    numerator = numerator * count / size        
+            numerator[user][item] = proportion[int(value / 0.2) - 1]
+    numerator = numerator * count / size
     _, _, ratings_less_biased, P, R, R_no_noise = ground_truth_matrix_to_dataset(
         truth, quantization='binary', bias='popularity', beta=beta, sample_prob=1)
     for key, value in ratings.items():
[…]

### 2022-08-07 algo.py
Finish drafting stage 2 experiments

@@ -52,7 +52,7 @@ class SVD(AlgoBase):
         r, err, dot, puf, qif = 0., 0., 0., 0., 0.
         global_mean = self.trainset.global_mean
 
-        lr_bu = self.lr_bucod
+        lr_bu = self.lr_bu
         lr_bi = self.lr_bi
         lr_pu = self.lr_pu
         lr_qi = self.lr_qi

### 2022-08-09 paper/paper.tex
Finish drafting stage 2 experiments
Run experiment

@@
 % ready for submission
-\usepackage{neurips_2022}
+% \usepackage{neurips_2022}
 
@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-%     \usepackage[final]{neurips_2022}
+\usepackage[preprint]{neurips_2022}
 
@@
 
-\title{Propensity Estimation Matrix Factorization Through Mixed-Quantized Dataset}
+\title{Data-Centric Recommender System for Tackling Selection Bias-induced Distribution Shift}
 
@@
 \author{%
-  David S.~Hippocampus\thanks{Use footnote for providing further information
-    about author (webpage, alternative address)---\emph{not} for acknowledging
-    funding agencies.} \\
-  Department of Computer Science\\
-  Cranberry-Lemon University\\
-  Pittsburgh, PA 15213 \\
-  \texttt{hippo@cs.cranberry-lemon.edu} \\
-  % examples of more authors
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \AND
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
-  % \And
-  % Coauthor \\
-  % Affiliation \\
-  % Address \\
-  % \texttt{email} \\
+  Fengyu Li \\
+  Cornell University\\
+  \texttt{fl334@cornell.edu} \\
+  \And
+  Sarah Dean \\
+  Cornell University \\
+  \texttt{sdean@cornell.edu} \\
 }

### 2022-08-09 data_utils.py

@@ -233,6 +233,9 @@ class MLP(nn.Module):
 
 
 def mlp_propensity_estimation(ratings):
+    '''
+    Not working yet.
+    '''
     X = []
     y = []
     for (user, item), value in ratings.items():

### 2022-08-10 algo.py
Run experiment

@@ -256,10 +256,10 @@ if __name__ == '__main__':
     data = Dataset.load_from_df(
         df[['userID', 'itemID', 'rating']], reader)
     trainset = data.build_full_trainset()
-    p = masked_nb_propensity_estimation(truth, ratings, P.shape)
+    p = masked_nb_propensity_estimation(truth, ratings, P.shape, beta=0)
 
-    algo_better = PropensitySVD(p, n_epochs=5, verbose=True)
-    algo = SVD(n_epochs=5, verbose=True)
+    algo_better = PropensitySVD(p, n_epochs=20, verbose=True)
+    algo = SVD(n_epochs=20, verbose=True)
     algo_better.fit(trainset)
     algo.fit(trainset)
     test_df = generate_test_dataframe(R_no_noise)

### 2022-08-10 data_utils.py

@@ -281,7 +281,7 @@ def masked_nb_propensity_estimation(truth, ratings, shape, beta=0):
         truth, quantization='binary', bias='popularity', beta=beta, sample_prob=1)
     for key, value in ratings.items():
         if value is not None and ratings_less_biased[key] is None:
-            value = None
+            ratings[key] = None
     proportion = np.zeros(5)
     for value in ratings.values():
         if value is not None:

### 2022-08-10 algo.py
Fix bug that causes wrong propensity estimation

@@ -4,7 +4,7 @@ from __future__ import (absolute_import, division, print_function,
 # import os
 # os.environ['KMP_DUPLICATE_LIB_OK']='True'
                         
-from data_utils import generate_ground_truth_matrix, ground_truth_matrix_to_dataset, to_dataframe, generate_test_dataframe, masked_nb_propensity_estimation
+from data_utils import generate_ground_truth_matrix, ground_truth_matrix_to_dataset, to_dataframe, generate_test_dataframe, masked_nb_propensity_estimation, naive_propensity_estimation
 import numpy as np
 from surprise import Reader
 from surprise import Dataset
@@ -250,18 +250,20 @@ if __name__ == '__main__':
     truth = generate_ground_truth_matrix(
         (1000, 1000), environment='ml-100k-v1')
     users, items, ratings, P, R, R_no_noise = ground_truth_matrix_to_dataset(
-        truth, quantization='onetofive', bias='popularity')
+        truth, quantization='onetofive', bias='full underlying', beta=1)
     df = to_dataframe(ratings)
     reader = Reader(rating_scale=(0, 1))
     data = Dataset.load_from_df(
         df[['userID', 'itemID', 'rating']], reader)
     trainset = data.build_full_trainset()
-    p = masked_nb_propensity_estimation(truth, ratings, P.shape, beta=0)
-
-    algo_better = PropensitySVD(p, n_epochs=20, verbose=True)
-    algo = SVD(n_epochs=20, verbose=True)
+    p = masked_nb_propensity_estimation(truth, ratings, P.shape, beta=0.5)
+    p_naive = naive_propensity_estimation(ratings, P.shape)
+    algo_better = PropensitySVD(p, n_epochs=5, verbose=True)
+    algo = PropensitySVD(p_naive, n_epochs=5, verbose=True)
+    algo_worst = SVD(n_epochs=5, verbose=True)
     algo_better.fit(trainset)
     algo.fit(trainset)
+    algo_worst.fit(trainset)
     test_df = generate_test_dataframe(R_no_noise)
     testset = Dataset.load_from_df(
         test_df[['userID', 'itemID', 'rating']], reader).build_full_trainset().build_testset()
@@ -270,4 +272,7 @@ if __name__ == '__main__':
     print(accuracy.rmse(predictions))
 
     predictions = algo.test(testset)
+    print(accuracy.rmse(predictions))
+
+    predictions = algo_worst.test(testset)
     print(accuracy.rmse(predictions))

### 2022-08-10 data_utils.py

@@ -7,6 +7,7 @@ from random import random
 from scipy.stats import pearsonr, spearmanr
 import pandas as pd
 import numpy as np
+from copy import deepcopy
 import os
 os.environ['KMP_DUPLICATE_LIB_OK'] = 'True'
 
@@ -210,7 +211,8 @@ def naive_propensity_estimation(ratings, shape, quantization='onetofive'):
     proportion /= proportion.sum()
     m, n = shape
     size = m * n
-    return np.zeros(shape) + count / size
+    result = np.zeros(shape) + count / size
+    return result / result.mean()
 
 
 class MLP(nn.Module):
@@ -277,22 +279,39 @@ def masked_nb_propensity_estimation(truth, ratings, shape, beta=0):
             user, item = key
             numerator[user][item] = proportion[int(value / 0.2) - 1]
     numerator = numerator * count / size
+
     _, _, ratings_less_biased, P, R, R_no_noise = ground_truth_matrix_to_dataset(
-        truth, quantization='binary', bias='popularity', beta=beta, sample_prob=1)
+        truth, quantization='binary', bias='full underlying', beta=beta, sample_prob=1)
+    ratings_copy = deepcopy(ratings)
+
     for key, value in ratings.items():
         if value is not None and ratings_less_biased[key] is None:
-            ratings[key] = None
+            ratings_copy[key] = None
+
     proportion = np.zeros(5)
-    for value in ratings.values():
+    for value in ratings_copy.values():
         if value is not None:
             proportion[int(value / 0.2) - 1] += 1
     proportion /= proportion.sum()
+
+    # just check
+    # _, _, ratings_ml100k, _, _, _ = ground_truth_matrix_to_dataset(
+    #     truth, quantization='onetofive', bias=None)
+    # proportion_truth = np.zeros(5)
+    # for value in ratings_ml100k.values():
+    #     if value is not None:
+    #         proportion_truth[int(value / 0.2) - 1] += 1
+    #         count += 1
+    # proportion_truth /= proportion_truth.sum()
+    # print(proportion_truth)
+    # print(proportion)
+
     denominator = np.ones(shape)
     for (user, item), value in ratings.items():
         if value is not None:
             denominator[user][item] = proportion[int(value / 0.2) - 1]
     result = np.divide(numerator, denominator)
-    result[result == 0] = 1
+    result[result == 0] = 0.1
     return result
 
 
@@ -300,9 +319,9 @@ if __name__ == '__main__':
[…]

### 2022-08-14 algo.py
RMSE vs epoch

@@ -12,6 +12,7 @@ from surprise import accuracy
 from surprise.prediction_algorithms import AlgoBase
 from surprise.prediction_algorithms import PredictionImpossible
 from surprise.utils import get_rng
+from copy import deepcopy
 
 class SVD(AlgoBase):
 
@@ -19,7 +20,7 @@ class SVD(AlgoBase):
                  init_std_dev=.1, lr_all=.005,
                  reg_all=.02, lr_bu=None, lr_bi=None, lr_pu=None, lr_qi=None,
                  reg_bu=None, reg_bi=None, reg_pu=None, reg_qi=None,
-                 random_state=None, verbose=False):
+                 random_state=None):
 
         self.n_factors = n_factors
         self.n_epochs = n_epochs
@@ -35,18 +36,21 @@ class SVD(AlgoBase):
         self.reg_pu = reg_pu if reg_pu is not None else reg_all
         self.reg_qi = reg_qi if reg_qi is not None else reg_all
         self.random_state = random_state
-        self.verbose = verbose
+        self.models = []
 
         AlgoBase.__init__(self)
 
-    def fit(self, trainset):
+    def fit(self, trainset, verbose=False):
 
         AlgoBase.fit(self, trainset)
-        self.sgd(trainset)
+        self.sgd(trainset, verbose=verbose)
 
-        return self
+        if not verbose:
+            return self
+        else:
+            return self, self.models
 
-    def sgd(self, trainset):
+    def sgd(self, trainset, verbose=False):
 
         u, i, f = 0, 0, 0
         r, err, dot, puf, qif = 0., 0., 0., 0., 0.
@@ -75,8 +79,8 @@ class SVD(AlgoBase):
             global_mean = 0
 
         for current_epoch in range(self.n_epochs):
-            if self.verbose:
-                print("Processing epoch {}".format(current_epoch))
+            # if self.verbose:
+            #     print("Processing epoch {}".format(current_epoch))
             for u, i, r in trainset.all_ratings():
 
                 # compute current error
@@ -97,10 +101,13 @@ class SVD(AlgoBase):
                     pu[u, f] += lr_pu * (err * qif - reg_pu * puf)
                     qi[i, f] += lr_qi * (err * puf - reg_qi * qif)
 
-        self.bu = bu
-        self.bi = bi
-        self.pu = pu
-        self.qi = qi
+            self.bu = bu
+            self.bi = bi
+            self.pu = pu
[…]

### 2022-08-17 paper/paper.tex
RMSE vs epoch
Continue paper on overleaf

@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-\usepackage[preprint]{neurips_2022}
+\usepackage{neurips_2022}
 
@@
 \usepackage{amsmath}
+\bibliographystyle{ksfh_nat}
 
@@
 
-\title{Data-Centric Recommender System for Tackling Selection Bias-induced Distribution Shift}
+\title{Cross-Dataset Recommender System for \\Overcoming Selection Bias}
 
@@
   bold, and in point size 12. Two line spaces precede the abstract. The abstract
-  must be limited to one paragraph.
+  must be limited to one paragraph. The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
+  both the left- and right-hand margins. Use 10~point type, with a vertical
+  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
+  bold, and in point size 12. Two line spaces precede the abstract. The abstract
+  must be limited to one paragraph. testf
 \end{abstract}
@@
 \section{Introduction}
-Selection bias is one of the most prevalent source of biases in a dataset for recommender systems.
-Selection bias happens when there is a pattern in the items that a user rates, either implicitly or explicitly.
-For example, in a recommender system for movies, users might only watch and rate the movies that are attractive to them in the first place.
-Selection bias is particularly relevant to recommender systems because when the system actively recommends items that the users like, it is creating bias by preventing users from accessing uncovered items.
-Selection bias introduces a Missing Not At Random (MNAR) condition to the observed dataset by conditioning it on the predictions of the recommender system.
+Selection bias is one of the most prevalent sources of distribution shift for recommender systems.
+Selection bias happens when there is a pattern in the users' ratings that is unique to the training set, either implicitly or explicitly.
+For example, in a recommender system for movies, users might only watch and rate movies that they particularly like, while distribution the recommender system is deployed on contains all movies regardless of person al tastes.
 Tackling the selection bias in a recommender dataset has been a constant challenge in designing recommender algorithms
@@
 \begin{enumerate}
-  \item We design experiments under a simulated environment and shows that a more quantized dataset is more susceptible to selection bias (and thus the MNAR condition.)
-  \item We propose a cross-dataset learning framework that empowers existing recommender algorithms to make use of the more quantized, implicit data. \textcolor{red}{EXPLAIN MORE HERE}
+  \item We design experiments under a simulated environment and shows that susceptibility to selection bias is likely not correlated with the way a dataset is quantized.
+  \item We propose a cross-dataset learning framework that empowers existing recommender algorithms to make use of the more quantized, less biased data.
+We carried out experiment and found our method outperforming baselines by a significant margin.
 \end{enumerate}
-
-Our hypothesis and theoretical results are verified through a series of empirical evaluation in a simulated environment where users' behaviors are carefully modeled.
-We show that our learning framework is significantly better than baseline recommender systems that neglects selection bias.
-We also show that our methods are substantially better than its single-dataset counterpart.
-\textcolor{red}{EXPLAIN MORE HERE}
 
@@
 
-\subsection{Recommender Systems Against Selection Bias}
+\section{Susceptibility to Selection Bias}
 
-\subsection{Quantization}
+\subsection{Simulated Environment for Controlling Bias}
 
-\section{Cross-Dataset Learning Framework}
+\subsection{Results}
+\bibliography{refs}
+
+\section{Cross-Dataset Propensity Estimation}
 
@@
 
-\subsection{Experiment Setup}
-
-\subsection{Controlling Bias}
+\subsection{Dataset and Baselines}
 
@@
 \section{Conclusion}
+
+\section*{Acknowledgements}
+
+\section*{References}
+\bibliography{refs}
 

### 2022-08-18 paper/paper.tex
Write introduction

@@
 \usepackage{neurips_2022}
+\bibliographystyle{plain}
 
@@
 \usepackage{amsmath}
-\bibliographystyle{ksfh_nat}
 
@@
   bold, and in point size 12. Two line spaces precede the abstract. The abstract
-  must be limited to one paragraph. testf
+  must be limited to one paragraph.
 \end{abstract}
@@
 \section{Introduction}
-Selection bias is one of the most prevalent sources of distribution shift for recommender systems.
-Selection bias happens when there is a pattern in the users' ratings that is unique to the training set, either implicitly or explicitly.
-For example, in a recommender system for movies, users might only watch and rate movies that they particularly like, while distribution the recommender system is deployed on contains all movies regardless of person al tastes.
-Tackling the selection bias in a recommender dataset has been a constant challenge in designing recommender algorithms
+Selection bias is one of the most prevalent sources of biases for recommender systems \cite{chen2020bias}.
+Selection bias happens when there is a pattern in the users' ratings that is unique to the training set.
+For example, in a recommender system for movies, users might mainly rate movies that are recommended to them, which is a small section of movies already tailored to the users' propensities.
+However, the environment the recommender system is deployed on contains all movies regardless of personal tastes.
+This discrepancy produces a misalignment between training and deployed settings, which is known as a distribution shift.
+Tackling the selection bias in a recommender dataset has been a constant challenge in designing recommender algorithms\cite{schnabel2016recommendations}.
 
-Quantization is a practical way to categorize an input from a continuous distribution to a discrete set of bins.
+When recommender systems are deployed in real-world platforms, it is arguably likely that the clients are able to collect rating-associated data from different source distributions.
+These data (or feedbacks) are either implicit or explicit \cite{aggarwal2016recommender}, while implicit data often comes in a highly-quantized, binary form and explicit data is usually less quantized.
+For example, a user's explicit rating of a movie is usually quantized to a number between 1 and 5, while the implicit feedback, such as if a user spontaneous searches for a movie, is often binary (more quantized.) We extend our hypothesis and argue that implicit data contains less or no selection bias compared to explicit data.
+This is because implicit data are often from users' spontaneous actions while explicit data are prejudiced toward the output of the recommender system in the previous feedback loop and users' innate biases.
+
+In this paper, we attempt to take advantage of datasets from differently quantized sources.
+More specifically, we proposed a way to feed both a 5-quantized dataset and a binary dataset to any gradient-based recommender algorithm.
+To ensure both datasets do not significantly lose their values in the presence of selection bias, we first examine the susceptibility to selection bias of differently quantized datasets from a single distribution.
+We design experiments under a simulated environment and shows that susceptibility to selection bias is not correlated with the way a dataset is quantized.
+
+Then, since a less-quantized dataset inherently contains more information than a more-quantized dataset \cite{widrow1996statistical} and thus is more suitable for training, we decided to use it as the training data of matrix factorization and use the more-quantized dataset for propensity scoring and deriving the inverse-probability-scoring (IPS) estimator \cite{thompson2012sampling} \cite{imbens2015causal}, a causal inference approach applicable to matrix completion-based recommender algorithms.
+In this way, our cross-dataset learning framework empowers existing recommender algorithms to make use of the more quantized, less biased data.
+We carried out experiment and found our method outperforming baselines by a significant margin.
+
+% Quantization is a practical way to categorize an input from a continuous distribution to a discrete set of bins.
 We assume the ground truth of a user's rating is a continuous distribution over items because of the inherent impossibility of numerically expressing the feeling of enjoyment and excitement.
@@
 
-We hypothesize that quantization and the effect of selection bias on the recommender system is positively correlated: the more quantized the dataset is (the fewer bins there are), the more severe the selection bias becomes in terms of the recommender system's performance.
-Therefore, a good-performing recommender system should prioritize training data that is less quantized.
-
-In reality, however, less quantized data might only represent a small subset of all the data one is capable of collecting.
-This is because less quantized data is typically explicit while more quantized data is typically implicit.
-That is to say, less quantized data would require a user explicitly rating an item, while more quantized data would be more easily gathered by observing the users' behaviors.
-For example, by observing the movie a user spontaneously chooses to see or if a user watches a movie's trailer to the end, we can easily create a 1-quantized or 2-quantized dataset, correspondingly, without the user paying any effort.
-In this paper, we propose a novel framework that utilizes these data.
-Our framework specifically exploits the fact that more quantized data is more susceptible to selection bias and treats it as an advantage that helps less quantized dataset better overcome the damage caused by selection bias.
-
-The main contribution of this work is two-fold:
-\begin{enumerate}
-  \item We design experiments under a simulated environment and shows that susceptibility to selection bias is likely not correlated with the way a dataset is quantized.
-  \item We propose a cross-dataset learning framework that empowers existing recommender algorithms to make use of the more quantized, less biased data.
-We carried out experiment and found our method outperforming baselines by a significant margin.
-\end{enumerate}
-
 \section{Related Work}
+Prior works on overcoming selection bias-induced distribution shift fall
 
@@
 \subsection{Results}
-\bibliography{refs}
 
@@
 
-\section*{References}
 \bibliography{refs}

### 2022-08-18 paper/paper.tex
Write related works

@@
 Selection bias happens when there is a pattern in the users' ratings that is unique to the training set.
-For example, in a recommender system for movies, users might mainly rate movies that are recommended to them, which is a small section of movies already tailored to the users' propensities.
+For example, in a recommender system for movies, users might mainly rate movies that are recommended to them, which is a small section of movies already tailored to the users' tastes \cite{pradel2012ranking}.
 However, the environment the recommender system is deployed on contains all movies regardless of personal tastes.
@@
 For example, a user's explicit rating of a movie is usually quantized to a number between 1 and 5, while the implicit feedback, such as if a user spontaneous searches for a movie, is often binary (more quantized.) We extend our hypothesis and argue that implicit data contains less or no selection bias compared to explicit data.
-This is because implicit data are often from users' spontaneous actions while explicit data are prejudiced toward the output of the recommender system in the previous feedback loop and users' innate biases.
+This is because implicit data are often from users' spontaneous actions while explicit data are prejudiced toward the output of the recommender system in the previous feedback loop and also users' innate biases.
 
@@
 \section{Related Work}
-Prior works on overcoming selection bias-induced distribution shift fall
+Prior works on overcoming selection bias-induced distribution shift via a propensity-based approach begins with the seminal paper \cite{schnabel2016recommendations}, which introduces the IPS method into recommender systems.
+Follow-up works aim at providing a learning-based or behavioral model of user feedbacks for propensity estimation \cite{joachims2017unbiased} \cite{yang2018unbiased}, which remains the central concern of this approach.
+In this work we add to previous attempts the idea of using a more quantized, implicit dataset for more accurate propensity estimation.
+
+The IPS method in our context is, essentially, a method of weighting training examples to correct the bias in the training data.
+Equivalent approaches such as importance weighting are widely used for domain adaptation in fields other than recommender systems \cite{sugiyama2007covariate} \cite{zhang2018importance}.
+Discussions on the IPS-based domain adaptation for countering selection bias, which is most relevant to recommender systems, remain limited.
 

### 2022-08-20 paper/paper.tex
Write 3.1

@@
 \section{Susceptibility to Selection Bias}
+We first examine the susceptibility to selection bias of differently quantized data by manually introducing biased distributions of various degrees to the differently quantized training sets.
+It is crucial for us that the 2-quantized datasets do not exhibit particular weakness when facing selection bias so that they can be properly adopted for propensity estimation.
 
 \subsection{Simulated Environment for Controlling Bias}
+Since selection bias is uncontrollable in a dataset completely drawn from real-world, we have to adopt a simulated environment \cite{krauth2020offline} with both semi-synthetic and synthetic datasets, which shall be explained in section 5.
+In our environment, we propose the \textbf{softmax observation model} and introduce a hyperparameter $\beta$ to control the degree of bias.
+For a rating matrix $R$, the corresponding probability matrix of each rating being observed is $\Pr(R_{u,i}\text{ is observed}) = k\text{softmax}(R_{u,i})$, where $k$ is set so that the expected proportion of observed ratings is controlled.
+The effect of $\beta$ on probability of being observed for different ratings is visualized in xxx.
+In our experiment we assume constantly $10\%$ of ratings are observed.
+\begin{figure}
+  \centering
+  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+  \caption{Sample figure caption.}
+\end{figure}
 
@@
 
-Stochastic gradient descent (SGD) step for one sample rating $(i,j)$:
-\[ b_u \gets b_u - \alpha(b_u - \frac{1}{P_{i,j}}(r_{i,j} - \hat{r}_{i,j})) \]
-\[ b_i \gets b_i - \alpha(b_i - \frac{1}{P_{i,j}}(r_{i,j} - \hat{r}_{i,j})) \]
-\[ V_i \gets V_i - \alpha(V_i - \frac{1}{P_{i,j}}W_j(r_{i,j} - \hat{r}_{i,j})) \]
-\[ W_j \gets W_j - \alpha(W_j - \frac{1}{P_{i,j}}V_i(r_{i,j} - \hat{r}_{i,j})) \]
-where $\alpha$ is the learning rate, which can be tuned independently for each parameter. In our experiments, we fix $\alpha$ across parameters.
-
 \section{Experiment}
 We designed experiments under simulated environments that verify the results introduced above.
-We first verify that more quantized dataset is more susceptible to selection bias and MNAR.
-We then runs our learning framework and compares it with multiple baselines.
 
-\subsection{Dataset and Baselines}
+\begin{table}
+  \caption{Test set RMSE and MAE for MF-NBPE and baselines}
+  \label{sample-table}
+  \centering
+  \begin{tabular}{lll}
+    \toprule
+    \multicolumn{2}{c}{Part}                   \\
+    \cmidrule(r){1-2}
+    Name     & Description     & Size ($\mu$m) \\
+    \midrule
+    Dendrite & Input terminal  & $\sim$100     \\
+    Axon     & Output terminal & $\sim$10      \\
+    Soma     & Cell body       & up to $10^6$  \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+\subsection{Datasets}
+We used two datasets in our simulated environment:\\
+
+\textbf{Imputed ML100K Dataset.}
+
+\textbf{Latent Factors Simulated Dataset.}
+
+\subsection{Baselines}
 
@@
 
-\section{Draft Area}
- (Consider using 2-dimensional Gaussian?)\\
-Assume the (flattened) ground truth ratings is a standard Gaussian with differential entropy
-\begin{align} \HH(R\sim N(0, 1))
-   & = \frac{1}{2}\log 2\pi e \\
-   & = 2.05
-\end{align}
-We can show that quantization (discretization) loses information. Consider
-a simple binary quantization that splits the ratings around the mean. Denote the quantized r.v. as $Q_1$.
-\[ \HH(Q_1) = -2 \cdot \frac{1}{2} \log \frac{1}{2} = 1 \]
-Consider another quantized r.v. $Q_2$ that divides the distribution four-fold based on the four quantiles.
-\[ \HH(Q_2) = -4 \cdot \frac{1}{4} \log \frac{1}{4} = 2 \]
-We can see that different quantization schemes loses different amount of information.
-\[ \HH(R) > \HH(Q_2) > \HH(Q_1) \]
-Next, we can prove that selection bias causes information loss.
-Assuming the selection bias causes ratings to have unequal probability to be sampled, and higher ratings are more likely to be sampled.
-The distribution of the observed ratings would then be right-skewed.
-We prove that, in the binary case, this reduces the information loss.
-Let $p$ be the probability that the sampled rating is positive.
-\[\HH = p\log p + (1-p) \log (1-p) \]
-\[ \frac{d\HH}{dp} = 1 + \log p + \frac{1-p}{p-1} - \log(1-p) = 0 \]
-which is optimized when $p = 0.5$.
-This result extends to all $n$-quantizations (can be proved using Lagrange multipliers.) See \url{http://pillowlab.princeton.edu/teaching/statneuro2018/slides/notes08_infotheory.pdf}.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2022-08-20 paper/paper.tex
Write 3.2

@@
 \usepackage{amsmath}
+\usepackage{graphicx}
+\graphicspath{ {./figures/} }
 
@@
 In our environment, we propose the \textbf{softmax observation model} and introduce a hyperparameter $\beta$ to control the degree of bias.
-For a rating matrix $R$, the corresponding probability matrix of each rating being observed is $\Pr(R_{u,i}\text{ is observed}) = k\text{softmax}(R_{u,i})$, where $k$ is set so that the expected proportion of observed ratings is controlled.
-The effect of $\beta$ on probability of being observed for different ratings is visualized in xxx.
+For a rating matrix $R$, the corresponding probability matrix of each rating being observed is $\Pr(R_{u,i}\text{ is observed}) = k\text{softmax}(\beta R_{u,i})$, where $k$ is set so that the expected proportion of observed ratings is controlled.
+The effect of $\beta$ on probability of being observed for different ratings is visualized in figure 1.
 In our experiment we assume constantly $10\%$ of ratings are observed.
@@
   \centering
-  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  \caption{Sample figure caption.}
+  \includegraphics[width=\textwidth]{visualize.png}
+  \caption{Visualizing the effect of controlling bias}
 \end{figure}
@@
 \subsection{Results}
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{1.png}
+  \caption{The effect of selection bias is not affected by quantization}
+\end{figure}
+Figure 2 shows our results on two dataset (see section 5) and 3 classic algorithms: user-KNN, item-KNN, and SVD matrix factorization.
+Although the RMSE grows consistently as $\beta$ increases for all datasets and algorithms, differently quantized datasets do not exhibit significantly different growth rates.
+We thus conclude it is a viable approach to use the more quantized, less biased dataset for propensity estimation.
 
@@
 We designed experiments under simulated environments that verify the results introduced above.
+
+\subsection{Datasets}
+We used two datasets in our simulated environment:\\
+
+\textbf{Imputed ML100K Dataset.}
+
+\textbf{Latent Factors Simulated Dataset.}
+
+\subsection{Baselines}
+
+\subsection{Results}
 
@@
 
-\subsection{Datasets}
-We used two datasets in our simulated environment:\\
-
-\textbf{Imputed ML100K Dataset.}
-
-\textbf{Latent Factors Simulated Dataset.}
-
-\subsection{Baselines}
-
-\subsection{Results}
-
 \section{Conclusion}

### 2022-08-22 paper/paper.tex
Write 5.1, 5.2

@@
 \section{Cross-Dataset Propensity Estimation}
+With the safety guarantee introduced above, we now turn to our cross-dataset matrix factorization model.
+Matrix factorization is a simple recommender model that decomposes the rating matrix based on known entries and then predicts the unknown entries\cite{koren2009matrix}.
+We integrate propensities as \cite{schnabel2016recommendations} did and formulated the recommendation problem as the empirical risk minimization framework below.
+\begin{equation} \argmin_{V,W,A} \frac{1}{N}\left(\frac{(Y_{u,i}- (V_u^TW_i + A))^2}{P_{u,i}} \right) + c \Vert A \Vert^2
+\end{equation}
+where $A = \{ b_u, b_i, \mu \}$ represents the standard bias parameters (offset), $V$ and $W$ are the decomposed vectors, $\hat{Y} = V_u^TW_i + A$ is the predicted rating, $N$ is the number of ratings, and $c \Vert A \Vert^2$ is the regularizer.
+The inverse propensity scores $\nicefrac{1}{P_{u,i}}$ are multiplied to each rating during learning, which analogous to re-weighting ratings based on their biases.
 
-\[ \argmin_{V,W,A} \frac{1}{N}\left(\frac{(Y_{u,i}- (V_u^TW_i + A))^2}{P_{u,i}} \right) + c \Vert A \Vert^2\]
-where $A = \{ b_u, b_i, \mu \}$ represents the standard bias parameters (offset), $\hat{Y} = V_u^TW_i + A$ is the predicted rating, and $c \Vert A \Vert^2$ is the regularizer.
+Denote the less quantized dataset for training as $D$. We propose the \textbf{naive-bayes propensity estimator} from a more quantized (binary) dataset $D'$. Essentially,
+\begin{equation} 
+  P_{u,i} = \Pr(Y_{u,i} \text{ is observed} \mid ) = \frac{}{\text{}},
+\end{equation}
+where xxx.
 
 \section{Experiment}
-We designed experiments under simulated environments that verify the results introduced above.
+We designed experiments to verify the performance of our cross-dataset model.
+We trained on two datasets and used two baseline algorithms for comparison.
+We selected the root mean square error (RMSE) and the mean absolute error (MAE) as the performance metrics.
 
 \subsection{Datasets}
-We used two datasets in our simulated environment:\\
-
 \textbf{Imputed ML100K Dataset.}
+The ML100K dataset provides 100 thousand MNAR ratings across 1683 movies rated by 944 users and is the standard large-scale dataset used for recommender systems.
+Since we need the ground truth ratings for controlling bias, we impute the missing ratings using standard matrix factorization.
 
 \textbf{Latent Factors Simulated Dataset.}
+The latent factors dataset is a synthetic dataset that models real-world user behavior.
+Users and item both have random latent vectors to simulate preferences; both also have biases.
+The environment is provided by \cite{krauth2020offline}.
 
 \subsection{Baselines}
+\textbf{Matrix Factorization.}
+As a simple baseline, we adopt the standard matrix factorization that does not adopt propensity estimation or importance weighting.
+
+\textbf{Naive Propensity Estimator.}
+The naive propensity estimator (NPE-MF) naively estimates the propensity scores from the already biased training data and plugs in the results to equation 1.
 
@@
 \begin{table}
-  \caption{Test set RMSE and MAE for MF-NBPE and baselines}
+  \caption{Test set RMSE and MAE for NBPE-MF and baselines}
   \label{sample-table}

### 2022-08-23 paper/paper.tex
Add a booktab table
Write section 4

@@
 Follow-up works aim at providing a learning-based or behavioral model of user feedbacks for propensity estimation \cite{joachims2017unbiased} \cite{yang2018unbiased}, which remains the central concern of this approach.
-In this work we add to previous attempts the idea of using a more quantized, implicit dataset for more accurate propensity estimation.
+While many works rely on a propensity matrix that is assumed to be known, TODO our work provides an idea of using a more quantized, implicit dataset for accurate propensity estimation.
 
@@
 
-Denote the less quantized dataset for training as $D$. We propose the \textbf{naive-bayes propensity estimator} from a more quantized (binary) dataset $D'$. Essentially,
+Denote the biased training set as $D$. we propose the \textbf{naive-bayes propensity estimator} from a more quantized (binary) dataset $D'$. Essentially,
 \begin{equation} 
-  P_{u,i} = \Pr(Y_{u,i} \text{ is observed} \mid ) = \frac{}{\text{}},
+  P_{u,i} = \Pr(Y_{u,i} \text{ is observed} \mid Y_{u,i} = r_{u,i}) = \frac{\Pr(Y_{u,i} = r_{u,i} \mid Y_{u,i} \text{ is observed})}{\Pr(Y_{u,i}) = r_{u,i}},
 \end{equation}
-where xxx.
+where the numerator can be easily approximated from the dataset, but the denominator requires additional data from less biased sources.
+In this paper, we estimates the denominator from the unbiased, more quantized dataset using a categorical naive-bayes model.
+We use $D'$ as a mask to hide biased ratings in the training set from propensity estimation.
+In other words, $Y_{u,i} \text{ is considered observed}$ only when $D'_{u,i} \text{ is observed}$.
+If the corresponding rating is not captured in $D'$, we set its propensity to the average propensity.
 
@@
 
+\textbf{Mixing Datasets MF.}
+Pending
+
 \subsection{Results}
@@
   \centering
-  \begin{tabular}{lll}
+  \begin{tabular}{lllll}
     \toprule               
-    \multicolumn{2}{c}{Part}                   \\
-    \cmidrule(r){1-2}
-    Name     & Description     & Size ($\mu$m) \\
+            & \multicolumn{2}{c}{ML100K} & \multicolumn{2}{c}{Latent Factors}              \\
+    \cmidrule(r){2-5}
+            & RMSE                       & MAE                                & RMSE & MAE \\
     \midrule
-    Dendrite & Input terminal  & $\sim$100     \\
-    Axon     & Output terminal & $\sim$10      \\
-    Soma     & Cell body       & up to $10^6$  \\
+    MF      & 0                          & 0                                  & 0    & 0   \\ 
+    NPE-MF  & 0                          & 0                                  & 0    & 0   \\ 
+    NBPE-MF & 0                          & 0                                  & 0    & 0   \\ 
     \bottomrule

### 2022-08-24 paper/paper.tex
Prepare for experiments
Finish experiment

@@
 Since we need the ground truth ratings for controlling bias, we impute the missing ratings using standard matrix factorization.
+Ratings are normalized between 0 and 1 for consistency with the other dataset.
 
@@
 \textbf{Mixing Datasets MF.}
-Pending
+To analyze whether our cross-dataset model more efficiently exploits all existing data, we create a larger dataset by mixing the two differently quantized datasets and train the MF algorithm on it.
 
@@
     \midrule
-    MF      & 0                          & 0                                  & 0    & 0   \\ 
-    NPE-MF  & 0                          & 0                                  & 0    & 0   \\ 
-    NBPE-MF & 0                          & 0                                  & 0    & 0   \\ 
+    MF               & 0.1046                     & 0.0833                             & 0.1331          & 0.1045          \\ 
+    NPE-MF           & 0.1048                     & 0.0834                             & 0.132           & 0.1049          \\ 
+    MD-MF            & 0.1044                     & 0.083                              & 0.1299          & 0.1027          \\
+    \textbf{NBPE-MF} & \textbf{0.076}             & \textbf{0.062}                     & \textbf{0.1065} & \textbf{0.0841} \\ 
     \bottomrule

### 2022-08-24 algo.py

@@ -4,12 +4,13 @@ from __future__ import (absolute_import, division, print_function,
 # import os
 # os.environ['KMP_DUPLICATE_LIB_OK']='True'
                         
-from data_utils import generate_ground_truth_matrix, ground_truth_matrix_to_dataset, to_dataframe, generate_test_dataframe, masked_nb_propensity_estimation, naive_propensity_estimation
+from data_utils import generate_ground_truth_matrix, ground_truth_matrix_to_dataset, to_dataframe, generate_test_dataframe, masked_nb_propensity_estimation, naive_propensity_estimation, mixing_mf_dataset
 import numpy as np
 from surprise import Reader
 from surprise import Dataset
 from surprise import accuracy
 from surprise.prediction_algorithms import AlgoBase
+from surprise.prediction_algorithms import SVD as surprise_SVD
 from surprise.prediction_algorithms import PredictionImpossible
 from surprise.utils import get_rng
 from copy import deepcopy
@@ -36,7 +37,6 @@ class SVD(AlgoBase):
         self.reg_pu = reg_pu if reg_pu is not None else reg_all
         self.reg_qi = reg_qi if reg_qi is not None else reg_all
         self.random_state = random_state
-        self.models = []
 
         AlgoBase.__init__(self)
 
@@ -45,10 +45,7 @@ class SVD(AlgoBase):
         AlgoBase.fit(self, trainset)
         self.sgd(trainset, verbose=verbose)
 
-        if not verbose:
-            return self
-        else:
-            return self, self.models
+        return self
 
     def sgd(self, trainset, verbose=False):
 
@@ -79,8 +76,8 @@ class SVD(AlgoBase):
             global_mean = 0
 
         for current_epoch in range(self.n_epochs):
-            # if self.verbose:
-            #     print("Processing epoch {}".format(current_epoch))
+            if verbose:
+                print("Processing epoch {}".format(current_epoch))
             for u, i, r in trainset.all_ratings():
 
                 # compute current error
@@ -105,9 +102,6 @@ class SVD(AlgoBase):
             self.bi = bi
             self.pu = pu
             self.qi = qi
-
-            if verbose:
-                self.models.append(deepcopy(self)) 
 
     def estimate(self, u, i):
         known_user = self.trainset.knows_user(u)
@@ -160,7 +154,6 @@ class PropensitySVD(AlgoBase):
[…]

### 2022-08-24 data_utils.py

@@ -145,6 +145,129 @@ def ground_truth_matrix_to_dataset(matrix, quantization, sample_prob=0.1, bias=N
         raise ValueError('Bias method not supported.')
 
 
+def mixing_mf_dataset(matrix, quantization, sample_prob=0.1, bias=None, beta=1, noise=0.05):
+    # Bias must be 'popularity', quantization must be 'onetofive'
+    m, n = matrix.shape
+    np.random.shuffle(matrix)
+    R = matrix.copy()
+
+    # Normalize the ground truth and add Gaussian noise
+    R_no_noise = (R - R.min()) / (R.max() - R.min())
+    R = R_no_noise + np.random.normal(0, noise, R.shape)
+    R[R > 1] = 1
+    R[R <= 0] = 1e-7
+    R_no_noise[R_no_noise <= 0] = 1e-7
+
+    # Quantize the matrix
+    if quantization == 'binary':
+        bins = np.linspace(0, 1, 3)
+        indexes = np.digitize(R, bins)
+        indexes_no_noise = np.digitize(R_no_noise, bins, right=True)
+        bins = np.append(bins, 1)
+        for i in range(m):
+            for j in range(n):
+                R[i][j] = bins[indexes[i][j]]
+                R_no_noise[i][j] = bins[indexes_no_noise[i][j]]
+    elif quantization == 'onetofive':
+        bins = np.linspace(0, 1, 6)
+        indexes = np.digitize(R, bins)
+        indexes_no_noise = np.digitize(R_no_noise, bins, right=True)
+        bins = np.append(bins, 1)
+        for i in range(m):
+            for j in range(n):
+                R[i][j] = bins[indexes[i][j]]
+                R_no_noise[i][j] = bins[indexes_no_noise[i][j]]
+    elif quantization == 'onetothree':
+        bins = np.linspace(0, 1, 4)
+        indexes = np.digitize(R, bins)
+        indexes_no_noise = np.digitize(R_no_noise, bins, right=True)
+        bins = np.append(bins, 1)
+        for i in range(m):
+            for j in range(n):
+                R[i][j] = bins[indexes[i][j]]
+                R_no_noise[i][j] = bins[indexes_no_noise[i][j]]
+    elif quantization == 'onetoten':
+        bins = np.linspace(0, 1, 11)
+        indexes = np.digitize(R, bins)
+        indexes_no_noise = np.digitize(R_no_noise, bins, right=True)
+        bins = np.append(bins, 1)
[…]

### 2022-08-25 paper/paper.tex
Add results

@@
 
-Denote the biased training set as $D$. we propose the \textbf{naive-bayes propensity estimator} from a more quantized (binary) dataset $D'$. Essentially,
+Denote the biased training set as $D$. we propose the \textbf{naive-bayes propensity estimator} (NBPE-MF) from a more quantized (binary) dataset $D'$. Essentially,
 \begin{equation} 
@@
 In other words, $Y_{u,i} \text{ is considered observed}$ only when $D'_{u,i} \text{ is observed}$.
+\begin{equation} 
+  NB equation here.
+\end{equation}
 If the corresponding rating is not captured in $D'$, we set its propensity to the average propensity.
@@
 We designed experiments to verify the performance of our cross-dataset model.
-We trained on two datasets and used two baseline algorithms for comparison.
+We trained on two datasets with $\beta =1$ in training set and used two baseline algorithms for comparison.
 We selected the root mean square error (RMSE) and the mean absolute error (MAE) as the performance metrics.
@@
 \subsection{Baselines}
-\textbf{Matrix Factorization.}
+\textbf{Matrix Factorization (MF.)}
 As a simple baseline, we adopt the standard matrix factorization that does not adopt propensity estimation or importance weighting.
 
-\textbf{Naive Propensity Estimator.}
-The naive propensity estimator (NPE-MF) naively estimates the propensity scores from the already biased training data and plugs in the results to equation 1.
+\textbf{Naive Propensity Estimator (NPE-MF.)}
+The naive propensity estimator naively estimates the propensity scores from the already biased training data and plugs in the results to equation 1.
 
-\textbf{Mixing Datasets MF.}
+\textbf{Mixing Datasets Matrix Factorization (MD-MF.)}
 To analyze whether our cross-dataset model more efficiently exploits all existing data, we create a larger dataset by mixing the two differently quantized datasets and train the MF algorithm on it.
@@
 \subsection{Results}
+Table 1 shows the the experiment's results, each entry as the average of five independent trials.
+Averaged across datasets and metrics, our approach gains 23.1\% reduction in error compared to matrix factorization, 23.1\% compared to naive propensity estimator, and 22.1\% compared to mixing datasets matrix factorization.
 
-\begin{table}
+Additionally, we observe that NPE-MF and MF have very similar performance across all settings.
+This suggests that a IPS-based model will not gain any performance boost if propensities are inaccurately estimated.
+Furthermore, we argue that our model is data-efficient, i.e., it beats the competitor model (MD-MF), demonstrating that utilizing less biased dataset for propensity estimation is a more viable approach than mixing it with other data, though the latter is sometimes a common engineering practice.
+On our benchmarks, mixing datasets MF achieves less than 2\% advantage compared to using only one dataset.
+
+We conclude that our cross-dataset model significantly outperforms baselines and provides a efficient way to make use of more quantized (implicit) data.
+
+\begin{table}0
   \caption{Test set RMSE and MAE for NBPE-MF and baselines}

### 2022-08-27 paper/paper.tex
Full draft

@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-\usepackage{neurips_2022}
+\usepackage[]{neurips_2022}
 \bibliographystyle{plain}
@@
 
-\title{Cross-Dataset Recommender System for \\Overcoming Selection Bias}
+\title{Cross-Dataset Propensity Estimation for \\Debiasing Recommender Systems}
 
@@
 \begin{abstract}
-  The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
-  both the left- and right-hand margins. Use 10~point type, with a vertical
-  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
-  bold, and in point size 12. Two line spaces precede the abstract. The abstract
-  must be limited to one paragraph. The abstract paragraph should be indented \nicefrac{1}{2}~inch (3~picas) on
-  both the left- and right-hand margins. Use 10~point type, with a vertical
-  spacing (leading) of 11~points.  The word \textbf{Abstract} must be centered,
-  bold, and in point size 12. Two line spaces precede the abstract. The abstract
-  must be limited to one paragraph.
+  Datasets for training recommender systems are often subject to severe distribution shift induced by users' selection biases.
+In this paper, we study the impact of selection bias on datasets with different quantization.
+We then leverage two differently quantized datasets from different source distributions to improve propensity estimation for the inverse probability scoring method from causal inference.
+Empirically, our approach gains significant performance improvement over single-dataset methods and alternative ways of combining two datasets.
 \end{abstract}
@@
 For example, in a recommender system for movies, users might mainly rate movies that are recommended to them, which is a small section of movies already tailored to the users' tastes \cite{pradel2012ranking}.
-However, the environment the recommender system is deployed on contains all movies regardless of personal tastes.
+However, the environment the recommender system is deployed on contains all candidates movies, which is not biased toward personal tastes.
 This discrepancy produces a misalignment between training and deployed settings, which is known as a distribution shift.
-Tackling the selection bias in a recommender dataset has been a constant challenge in designing recommender algorithms\cite{schnabel2016recommendations}.
+Tackling the selection bias in a recommender dataset has been a constant challenge in designing recommender algorithms \cite{schnabel2016recommendations}.
 
-When recommender systems are deployed in real-world platforms, it is arguably likely that the clients are able to collect rating-associated data from different source distributions.
-These data (or feedbacks) are either implicit or explicit \cite{aggarwal2016recommender}, while implicit data often comes in a highly-quantized, binary form and explicit data is usually less quantized.
-For example, a user's explicit rating of a movie is usually quantized to a number between 1 and 5, while the implicit feedback, such as if a user spontaneous searches for a movie, is often binary (more quantized.) We extend our hypothesis and argue that implicit data contains less or no selection bias compared to explicit data.
-This is because implicit data are often from users' spontaneous actions while explicit data are prejudiced toward the output of the recommender system in the previous feedback loop and also users' innate biases.
+When recommender systems are deployed in real-world platforms, clients are usually able to collect preference-associated data from different source distributions.
+These data, or feedbacks, are either implicit or explicit \cite{aggarwal2016recommender}, highly-quantized (e.g.
+binary) or less quantized (e.g.
+one to five stars.) However, state-of-the-art recommender systems tend to only train on one dataset, which is often less quantized.
+We hypothesize that combing more than one datasets with different degrees of bias in the training procedure would make the model more robust against selection bias.
 
-In this paper, we attempt to take advantage of datasets from differently quantized sources.
-More specifically, we proposed a way to feed both a 5-quantized dataset and a binary dataset to any gradient-based recommender algorithm.
-To ensure both datasets do not significantly lose their values in the presence of selection bias, we first examine the susceptibility to selection bias of differently quantized datasets from a single distribution.
+In this paper, we attempt to take advantage of datasets from differently biased sources with different quantization.
+More specifically, we propose a way to feed both a highly quantized and a less quantized dataset to a gradient-based recommender algorithm, given the assumption that the highly quantized dataset contains less bias.
+Practically, this assumption stands true under many settings.
+Being less biased toward personal taste has been an important reason that implicit dataset, which is often more quantized, is adopted for modeling \cite{amatriain2009like} \cite{jawaheer2010comparison}.
+
+To ensure both datasets do not significantly lose their values in the presence of selection bias, we first examine the susceptibility to selection bias of differently quantized datasets from a source distribution.
 We design experiments under a simulated environment and shows that susceptibility to selection bias is not correlated with the way a dataset is quantized.
 
-Then, since a less-quantized dataset inherently contains more information than a more-quantized dataset \cite{widrow1996statistical} and thus is more suitable for training, we decided to use it as the training data of matrix factorization and use the more-quantized dataset for propensity scoring and deriving the inverse-probability-scoring (IPS) estimator \cite{thompson2012sampling} \cite{imbens2015causal}, a causal inference approach applicable to matrix completion-based recommender algorithms.
-In this way, our cross-dataset learning framework empowers existing recommender algorithms to make use of the more quantized, less biased data.
+Then, since a less-quantized dataset inherently contains more information than a more-quantized dataset \cite{widrow1996statistical} and thus is more suitable for training the model, we decide to use it as the training data of matrix factorization and use the more quantized dataset for propensity scoring and deriving the inverse-probability-scoring (IPS) estimator \cite{thompson2012sampling} \cite{imbens2015causal}, a causal inference approach applicable to matrix completion-based recommender algorithms.
+In this way, our cross-dataset learning framework empowers existing recommender algorithms by using previously neglected, highly quantized dataset.
 We carried out experiment and found our method outperforming baselines by a significant margin.
@@
 Follow-up works aim at providing a learning-based or behavioral model of user feedbacks for propensity estimation \cite{joachims2017unbiased} \cite{yang2018unbiased}, which remains the central concern of this approach.
-While many works rely on a propensity matrix that is assumed to be known, TODO our work provides an idea of using a more quantized, implicit dataset for accurate propensity estimation.
+While many works rely on a propensity matrix that is assumed to be known, e.g.
+\cite{something}, our work provides an idea of using a more quantized, implicit dataset for accurate propensity estimation.
 
@@
 We first examine the susceptibility to selection bias of differently quantized data by manually introducing biased distributions of various degrees to the differently quantized training sets.
-It is crucial for us that the 2-quantized datasets do not exhibit particular weakness when facing selection bias so that they can be properly adopted for propensity estimation.
+It is crucial for us that the highly quantized dataset does not exhibit particular weakness when facing selection bias so that they can be properly adopted for propensity estimation.
 
@@
 In our environment, we propose the \textbf{softmax observation model} and introduce a hyperparameter $\beta$ to control the degree of bias.
-For a rating matrix $R$, the corresponding probability matrix of each rating being observed is $\Pr(R_{u,i}\text{ is observed}) = k\text{softmax}(\beta R_{u,i})$, where $k$ is set so that the expected proportion of observed ratings is controlled.
+For a rating matrix $R$, the corresponding probability matrix of each rating being observed is $\Pr(R_{u,i}\text{ is observed}) = k\text{softmax}(\beta R_{u,i})$, where $k$ is pre-set so that the expected proportion of observed ratings is controlled.
 The effect of $\beta$ on probability of being observed for different ratings is visualized in figure 1.
-In our experiment we assume constantly $10\%$ of ratings are observed.
+In our experiment we assume constantly $10\%$ of ratings are observed among $1000$ users and $1000$ items.
 \begin{figure}
@@
 \end{figure}
-Figure 2 shows our results on two dataset (see section 5) and 3 classic algorithms: user-KNN, item-KNN, and SVD matrix factorization.
-Although the RMSE grows consistently as $\beta$ increases for all datasets and algorithms, differently quantized datasets do not exhibit significantly different growth rates.
+Figure 2 shows our results on a synthetic dataset (see section 5) and 3 classic algorithms: user-KNN, item-KNN, and SVD matrix factorization.
+The three differently quantized variants come from the same source distribution (the dataset).
+Gaussian noise is introduced for simulating real-world user noise.
+Shaded regions denote the regions between quartiles.
+
+Although the RMSE grows consistently as $\beta$ increases for algorithms, differently quantized datasets do not exhibit significantly different growth rates.
+The discrepancy between the RMSE of differently quantized datasets with the same $\beta$ can be ascribed to the inherent information loss from high quantization \cite{widrow1996statistical}.
 We thus conclude it is a viable approach to use the more quantized, less biased dataset for propensity estimation.
+Furthermore, figure 2.
+visualizes the damaging effect of selection bias.
+The less quantized datasets breaks even with the most coarse binary dataset at $\beta=2.5$ versus $\beta=0$, suggesting the practical importance of debiasing recommender systems.
+The above experiments are repeated under a semi-synthetic environment and the conclusion holds.
 
 \section{Cross-Dataset Propensity Estimation}
-With the safety guarantee introduced above, we now turn to our cross-dataset matrix factorization model.
-Matrix factorization is a simple recommender model that decomposes the rating matrix based on known entries and then predicts the unknown entries\cite{koren2009matrix}.
+With the above safety guarantee, we now turn to our cross-dataset matrix factorization model.
+Matrix factorization is a simple recommender model that decomposes the rating matrix based on known entries and then predicts the unknown entries \cite{koren2009matrix}.
 We integrate propensities as \cite{schnabel2016recommendations} did and formulated the recommendation problem as the empirical risk minimization framework below.
@@
 where $A = \{ b_u, b_i, \mu \}$ represents the standard bias parameters (offset), $V$ and $W$ are the decomposed vectors, $\hat{Y} = V_u^TW_i + A$ is the predicted rating, $N$ is the number of ratings, and $c \Vert A \Vert^2$ is the regularizer.
-The inverse propensity scores $\nicefrac{1}{P_{u,i}}$ are multiplied to each rating during learning, which analogous to re-weighting ratings based on their biases.
+The inverse propensity scores $\nicefrac{1}{P_{u,i}}$ are multiplied to each rating during learning, which is analogous to re-weighting ratings based on their biases.
 
@@
 \begin{equation} 
-  P_{u,i} = \Pr(Y_{u,i} \text{ is observed} \mid Y_{u,i} = r_{u,i}) = \frac{\Pr(Y_{u,i} = r_{u,i} \mid Y_{u,i} \text{ is observed})}{\Pr(Y_{u,i}) = r_{u,i}},
+  P_{u,i} = \Pr(Y_{u,i} \text{ is observed} \mid Y_{u,i} = r_{u,i}) = \frac{\Pr(Y_{u,i} = r_{u,i} \mid Y_{u,i} \text{ is observed})}{\Pr(Y_{u,i} = r_{u,i})},
 \end{equation}
-where the numerator can be easily approximated from the dataset, but the denominator requires additional data from less biased sources.
-In this paper, we estimates the denominator from the unbiased, more quantized dataset using a categorical naive-bayes model.
+where $r_{u,i}$ denotes the value of a rating.
+The numerator can be easily approximated from the dataset, but the denominator requires additional data from less biased sources.
+In this paper, we estimates the denominator from the less biased dataset using a categorical naive-bayes model.
 We use $D'$ as a mask to hide biased ratings in the training set from propensity estimation.
-In other words, $Y_{u,i} \text{ is considered observed}$ only when $D'_{u,i} \text{ is observed}$.
+In other words, $Y_{u,i} \text{ is considered observed}$ only when $D'_{u,i} \text{ is observed}$:
 \begin{equation} 
-  NB equation here.
+  \Pr(Y_{u,i} = r_{u,i}) = \frac{\sum_1^N \vect 1_{(Y^D_{u,i} = r_{u,i})} \cdot \vect 1_{(Y^{D'}_{u,i} \text{ is observed})} }{N}.
 \end{equation}
@@
 We designed experiments to verify the performance of our cross-dataset model.
-We trained on two datasets with $\beta =1$ in training set and used two baseline algorithms for comparison.
-We selected the root mean square error (RMSE) and the mean absolute error (MAE) as the performance metrics.
+We trained on two datasets with $\beta =1$ in training set and used three baseline algorithms for comparison.
+In this paper we assumed the highly quantized dataset is unbiased with $\beta=0$, which is probable when the implicit dataset reflects users' spontaneous choices.
+Theoretically, our framework will work as long as the highly quantized dataset is less biased.
+We selected the root mean square error (RMSE) and the mean absolute error (MAE) as performance metrics.
 
@@
 Table 1 shows the the experiment's results, each entry as the average of five independent trials.
-Averaged across datasets and metrics, our approach gains 23.1\% reduction in error compared to matrix factorization, 23.1\% compared to naive propensity estimator, and 22.1\% compared to mixing datasets matrix factorization.
+Averaged across datasets and metrics, our approach gains $23.1\%$ reduction in error compared to matrix factorization, $23.1\%$ compared to naive propensity estimator, and $22.1\%$ compared to mixing datasets matrix factorization.
 
@@
 Furthermore, we argue that our model is data-efficient, i.e., it beats the competitor model (MD-MF), demonstrating that utilizing less biased dataset for propensity estimation is a more viable approach than mixing it with other data, though the latter is sometimes a common engineering practice.
-On our benchmarks, mixing datasets MF achieves less than 2\% advantage compared to using only one dataset.
+We nudge the way two datasets are mixed in MD-MF and discover similar results.
+On our benchmarks, mixing datasets MF consistently achieves less than $2\%$ advantage compared to using only one dataset.
 
@@
 
-\begin{table}0
+\begin{table}
   \caption{Test set RMSE and MAE for NBPE-MF and baselines}
@@
     MD-MF            & 0.1044                     & 0.083                              & 0.1299          & 0.1027          \\
-    \textbf{NBPE-MF} & \textbf{0.076}             & \textbf{0.062}                     & \textbf{0.1065} & \textbf{0.0841} \\ 
+    \textbf{NBPE-MF}& \textbf{0.076}             & \textbf{0.062}                     & \textbf{0.1065} & \textbf{0.0841} \\ 
     \bottomrule
@@
 \section{Conclusion}
+We propose an effective and data-efficient method of combing differently sourced datasets for training recommender systems.
+Our approach provides yet another way of accurate propensity estimation and explores a new potential of cross-dataset joint learning.
+We believe real-world recommender systems will benefit from data-centric frameworks like ours.
 
 \section*{Acknowledgements}
+This work is supported in part by a Cornell CSURP Grant,  {\color{blue}{\textit{is there some source of funding you'd like to acknowledge?}}}
 

### 2022-08-27 algo.py

@@ -262,14 +262,14 @@ if __name__ == '__main__':
     experiment = 2
 
     if experiment == 1:
-        p = masked_nb_propensity_estimation(truth, ratings, P.shape)
+        p = masked_nb_propensity_estimation(truth, ratings, P.shape, beta=3)
         p_naive = naive_propensity_estimation(ratings, P.shape)
         algo_better = PropensitySVD(p, n_epochs=10)
         algo = PropensitySVD(p_naive, n_epochs=10)
         algo_worst = SVD(n_epochs=10)
         algo_better.fit(trainset, verbose=True)
-        algo.fit(trainset, verbose=True)
-        algo_worst.fit(trainset, verbose=True)
+        # algo.fit(trainset, verbose=True)
+        # algo_worst.fit(trainset, verbose=True)
         test_df = generate_test_dataframe(R_no_noise)
         testset = Dataset.load_from_df(
             test_df[['userID', 'itemID', 'rating']], reader).build_full_trainset().build_testset()
@@ -278,13 +278,13 @@ if __name__ == '__main__':
         print(accuracy.rmse(predictions))
         print(accuracy.mae(predictions))
 
-        predictions = algo.test(testset)
-        print(accuracy.rmse(predictions))
-        print(accuracy.mae(predictions))
+        # predictions = algo.test(testset)
+        # print(accuracy.rmse(predictions))
+        # print(accuracy.mae(predictions))
 
-        predictions = algo_worst.test(testset)
-        print(accuracy.rmse(predictions))
-        print(accuracy.mae(predictions))
+        # predictions = algo_worst.test(testset)
+        # print(accuracy.rmse(predictions))
+        # print(accuracy.mae(predictions))
 
     elif experiment == 2:
         algo = surprise_SVD(n_epochs=10, verbose=True)    

### 2022-08-27 data_utils.py

@@ -228,7 +228,7 @@ def mixing_mf_dataset(matrix, quantization, sample_prob=0.1, bias=None, beta=1, 
         for i in range(m):
             for j in range(n):
                 if R[m][n] is None:
-                    R[m][n] = 0.2 if matrix[m][n] <= 0.5 else 1
+                    R[m][n] = 0.4 if matrix[m][n] <= 0.5 else 0.8
                     ratings[(m, n)] = R[m][n]
         return users, items, ratings, P, R, R_no_noise
 
@@ -404,8 +404,20 @@ def masked_nb_propensity_estimation(truth, ratings, shape, beta=0):
     numerator = numerator * count / size
 
     _, _, ratings_less_biased, P, R, R_no_noise = ground_truth_matrix_to_dataset(
-        truth, quantization='binary', bias='full underlying', beta=beta, sample_prob=1)
+        truth, quantization='binary', bias='full underlying', beta=beta, sample_prob=0.1)
     ratings_copy = deepcopy(ratings)
+    
+    # n0 = 0 
+    # n1 = 1
+    # for v in ratings_less_biased.values():
+    #     if v = 0.5:
+    #         n0 += 1
+    #     elif v == 1:
+    #         n1 += 1
+    #     elif v is not None:
+    #         print(v)
+    # print(n0)
+    # print(n1)                
 
     for key, value in ratings.items():
         if value is not None and ratings_less_biased[key] is None:

### 2022-09-19 paper/paper.tex
Edit paper

@@
 \begin{abstract}
-  Datasets for training recommender systems are often subject to severe distribution shift induced by users' selection biases.
+  Datasets for training recommender systems are often subject to distribution shift induced by users' and recommenders' selection biases.
 In this paper, we study the impact of selection bias on datasets with different quantization.
-We then leverage two differently quantized datasets from different source distributions to improve propensity estimation for the inverse probability scoring method from causal inference.
+We then leverage two differently quantized datasets from different source distributions to mitigate distribution shift by applying the inverse probability scoring method from causal inference.
 Empirically, our approach gains significant performance improvement over single-dataset methods and alternative ways of combining two datasets.
@@
 binary) or less quantized (e.g.
-one to five stars.) However, state-of-the-art recommender systems tend to only train on one dataset, which is often less quantized.
+one to five stars.) However, state-of-the-art recommender algorithms tend to focus on one dataset.
 We hypothesize that combing more than one datasets with different degrees of bias in the training procedure would make the model more robust against selection bias.
@@
 To ensure both datasets do not significantly lose their values in the presence of selection bias, we first examine the susceptibility to selection bias of differently quantized datasets from a source distribution.
-We design experiments under a simulated environment and shows that susceptibility to selection bias is not correlated with the way a dataset is quantized.
+Our experiment shows that susceptibility to selection bias is not correlated with the way a dataset is quantized.
 
-Then, since a less-quantized dataset inherently contains more information than a more-quantized dataset \cite{widrow1996statistical} and thus is more suitable for training the model, we decide to use it as the training data of matrix factorization and use the more quantized dataset for propensity scoring and deriving the inverse-probability-scoring (IPS) estimator \cite{thompson2012sampling} \cite{imbens2015causal}, a causal inference approach applicable to matrix completion-based recommender algorithms.
-In this way, our cross-dataset learning framework empowers existing recommender algorithms by using previously neglected, highly quantized dataset.
+Then, since a less-quantized dataset inherently contains more information than a more-quantized dataset \cite{widrow1996statistical}, we decide to use it as the training set of matrix factorization and use the more quantized dataset for deriving the inverse-probability-scoring (IPS) propensity estimator \cite{thompson2012sampling} \cite{imbens2015causal}, a causal inference approach applicable to matrix completion-based recommender algorithms.
+In this way, our cross-dataset learning framework empowers existing recommender algorithms by empowering highly quantized dataset.
 We carried out experiment and found our method outperforming baselines by a significant margin.
@@
 Equivalent approaches such as importance weighting are widely used for domain adaptation in fields other than recommender systems \cite{sugiyama2007covariate} \cite{zhang2018importance}.
-Discussions on the IPS-based domain adaptation for countering selection bias, which is most relevant to recommender systems, remain limited.
+Discussions on the IPS-based domain adaptation for countering selection bias, which is most relevant to recommender systems, remain limited due to the hardness of evaluating propensities.
 
@@
 \subsection{Simulated Environment for Controlling Bias}
-Since selection bias is uncontrollable in a dataset completely drawn from real-world, we have to adopt a simulated environment \cite{krauth2020offline} with both semi-synthetic and synthetic datasets, which shall be explained in section 5.
+Since selection bias is uncontrollable in a dataset completely drawn from real-world, we have to adopt a simulated environment \cite{krauth2020offline} with both semi-synthetic and synthetic datasets explained below.
 In our environment, we propose the \textbf{softmax observation model} and introduce a hyperparameter $\beta$ to control the degree of bias.
 For a rating matrix $R$, the corresponding probability matrix of each rating being observed is $\Pr(R_{u,i}\text{ is observed}) = k\text{softmax}(\beta R_{u,i})$, where $k$ is pre-set so that the expected proportion of observed ratings is controlled.
-The effect of $\beta$ on probability of being observed for different ratings is visualized in figure 1.
+In this observation model, a larger $\beta$ induces more significant bias at an exponential rate.
 In our experiment we assume constantly $10\%$ of ratings are observed among $1000$ users and $1000$ items.
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{visualize.png}
-  \caption{Visualizing the effect of controlling bias}
-\end{figure}
+% \begin{figure}
+%   \centering
+%   \includegraphics[width=\textwidth]{visualize.png}
+%   \caption{Visualizing the effect of controlling bias}
+% \end{figure}
+
+\subsection{Datasets}
+\textbf{Imputed ML100K Dataset.}
+The ML100K dataset provides 100 thousand MNAR ratings across 1683 movies rated by 944 users and is the standard large-scale dataset used for recommender systems.
+Since we need the ground truth ratings for controlling bias, we impute the missing ratings using standard matrix factorization \cite{krauth2020offline}.
+Ratings are normalized between 0 and 1 for consistency with the other dataset.
+
+\textbf{Latent Factors Simulated Dataset.}
+The latent factors dataset is a synthetic dataset that models real-world user behavior.
+Users and item both have Gaussian latent vectors to simulate preferences \cite{koren2008factorization} \cite{koren2009matrix}; both also have biases.
+The environment is provided by \cite{krauth2020offline}.
 
@@
 \end{figure}
-Figure 2 shows our results on a synthetic dataset (see section 5) and 3 classic algorithms: user-KNN, item-KNN, and SVD matrix factorization.
+Figure 1 shows our results on the latent factors dataset with 3 classic algorithms: user-KNN, item-KNN, and SVD matrix factorization.
 The three differently quantized variants come from the same source distribution (the dataset).
@@
 
-Although the RMSE grows consistently as $\beta$ increases for algorithms, differently quantized datasets do not exhibit significantly different growth rates.
-The discrepancy between the RMSE of differently quantized datasets with the same $\beta$ can be ascribed to the inherent information loss from high quantization \cite{widrow1996statistical}.
+Although the RMSE grows consistently as $\beta$ increases, differently quantized datasets do not exhibit significantly different growth rates.
+The discrepancy between the RMSE of differently quantized datasets with the same $\beta$ can be ascribed to the inherent information loss from quantization \cite{widrow1996statistical}.
+We reason that since the effect of bias on rating predictions is consistent across quantizations, the subsequent effect on propensity estimation should be identical as accurate propensity estimation follows accurate rating predictions.
+Therefore, differently quantized datasets should also be equally susceptible to selection bias when used for propensity scoring.
 We thus conclude it is a viable approach to use the more quantized, less biased dataset for propensity estimation.
-Furthermore, figure 2.
+Furthermore, figure 1.
 visualizes the damaging effect of selection bias.
@@
 \section{Cross-Dataset Propensity Estimation}
-With the above safety guarantee, we now turn to our cross-dataset matrix factorization model.
+With the above guarantee, we now turn to our cross-dataset matrix factorization model.
 Matrix factorization is a simple recommender model that decomposes the rating matrix based on known entries and then predicts the unknown entries \cite{koren2009matrix}.
-We integrate propensities as \cite{schnabel2016recommendations} did and formulated the recommendation problem as the empirical risk minimization framework below.
+We integrate propensities as in \citep{schnabel2016recommendations} and formulated the recommendation problem as the empirical risk minimization framework below.
 \begin{equation} \argmin_{V,W,A} \frac{1}{N}\left(\frac{(Y_{u,i}- (V_u^TW_i + A))^2}{P_{u,i}} \right) + c \Vert A \Vert^2
@@
 
-\subsection{Datasets}
-\textbf{Imputed ML100K Dataset.}
-The ML100K dataset provides 100 thousand MNAR ratings across 1683 movies rated by 944 users and is the standard large-scale dataset used for recommender systems.
-Since we need the ground truth ratings for controlling bias, we impute the missing ratings using standard matrix factorization.
-Ratings are normalized between 0 and 1 for consistency with the other dataset.
-
-\textbf{Latent Factors Simulated Dataset.}
-The latent factors dataset is a synthetic dataset that models real-world user behavior.
-Users and item both have random latent vectors to simulate preferences; both also have biases.
-The environment is provided by \cite{krauth2020offline}.
-
 \subsection{Baselines}
@@
 To analyze whether our cross-dataset model more efficiently exploits all existing data, we create a larger dataset by mixing the two differently quantized datasets and train the MF algorithm on it.
+To do this, we tried multiple schemes to upscale the more quantized dataset to be less quantized and found that this has minimal influence on the results.
 
@@
     MD-MF            & 0.1044                     & 0.083                              & 0.1299          & 0.1027          \\
-    \textbf{NBPE-MF}& \textbf{0.076}             & \textbf{0.062}                     & \textbf{0.1065} & \textbf{0.0841} \\ 
+    \textbf{NBPE-MF} & \textbf{0.076}             & \textbf{0.062}                     & \textbf{0.1065} & \textbf{0.0841} \\ 
     \bottomrule
@@
 Our approach provides yet another way of accurate propensity estimation and explores a new potential of cross-dataset joint learning.
-We believe real-world recommender systems will benefit from data-centric frameworks like ours.
+We demonstrate that our method outperforms naive propensity estimation and straightforward dataset mixing.
+and  We believe real-world recommender systems will benefit from data-centric frameworks like ours.
 

### 2022-09-20 paper/paper.tex
Edit paper
Edit

@@
 We designed experiments to verify the performance of our cross-dataset model.
-We trained on two datasets with $\beta =1$ in training set and used three baseline algorithms for comparison.
+We trained on two datasets (as in section 3) with $\beta =1$ in training set and used three baseline algorithms for comparison.
 In this paper we assumed the highly quantized dataset is unbiased with $\beta=0$, which is probable when the implicit dataset reflects users' spontaneous choices.
-Theoretically, our framework will work as long as the highly quantized dataset is less biased.
+Empirically, our framework will work as long as the highly quantized dataset is less biased.
 We selected the root mean square error (RMSE) and the mean absolute error (MAE) as performance metrics.
@@
 
-\section*{Acknowledgements}
-This work is supported in part by a Cornell CSURP Grant,  {\color{blue}{\textit{is there some source of funding you'd like to acknowledge?}}}
+% \section*{Acknowledgements}
+% This work is supported in part by a Cornell CSURP Grant,  {\color{blue}{\textit{is there some source of funding you'd like to acknowledge?}}}
 

### 2022-09-26 paper/paper.tex
small edits and some comments in blueg

@@
 binary) or less quantized (e.g.
-one to five stars.) However, state-of-the-art recommender algorithms tend to focus on one dataset.
+one to five stars).
+However, state-of-the-art recommender algorithms tend to focus on one dataset.
 We hypothesize that combing more than one datasets with different degrees of bias in the training procedure would make the model more robust against selection bias.
 
-In this paper, we attempt to take advantage of datasets from differently biased sources with different quantization.
+In this paper, we attempt to take advantage of datasets from differently biased sources with different levels of quantization.
 More specifically, we propose a way to feed both a highly quantized and a less quantized dataset to a gradient-based recommender algorithm, given the assumption that the highly quantized dataset contains less bias.
 Practically, this assumption stands true under many settings.
-Being less biased toward personal taste has been an important reason that implicit dataset, which is often more quantized, is adopted for modeling \cite{amatriain2009like} \cite{jawaheer2010comparison}.
+Implicit datasets, which are usually more quantized, are often adopted because they are observed to be less biased towards personal taste \cite{amatriain2009like,jawaheer2010comparison}.
+% Being less biased toward personal taste has been an important reason that implicit dataset, which is often more quantized, is adopted for modeling \cite{amatriain2009like} \cite{jawaheer2010comparison}.
 
-To ensure both datasets do not significantly lose their values in the presence of selection bias, we first examine the susceptibility to selection bias of differently quantized datasets from a source distribution.
+To ensure both datasets do not significantly lose their value in the presence of selection bias, we first examine the susceptibility to selection bias of differently quantized datasets from a source distribution.
 Our experiment shows that susceptibility to selection bias is not correlated with the way a dataset is quantized.
 
-Then, since a less-quantized dataset inherently contains more information than a more-quantized dataset \cite{widrow1996statistical}, we decide to use it as the training set of matrix factorization and use the more quantized dataset for deriving the inverse-probability-scoring (IPS) propensity estimator \cite{thompson2012sampling} \cite{imbens2015causal}, a causal inference approach applicable to matrix completion-based recommender algorithms.
+Then, since a less-quantized dataset inherently contains more information than a more-quantized dataset \cite{widrow1996statistical}, we decide to use it as the training set of matrix factorization and use the more quantized dataset for deriving the inverse-probability-scoring (IPS) propensity estimator \cite{thompson2012sampling,imbens2015causal}, a causal inference approach applicable to matrix completion-based recommender algorithms.
 In this way, our cross-dataset learning framework empowers existing recommender algorithms by empowering highly quantized dataset.
-We carried out experiment and found our method outperforming baselines by a significant margin.
+We carried out an experiment and found our method to outperform baselines by a significant margin.
 
@@
 Prior works on overcoming selection bias-induced distribution shift via a propensity-based approach begins with the seminal paper \cite{schnabel2016recommendations}, which introduces the IPS method into recommender systems.
-Follow-up works aim at providing a learning-based or behavioral model of user feedbacks for propensity estimation \cite{joachims2017unbiased} \cite{yang2018unbiased}, which remains the central concern of this approach.
+Follow-up works aim at providing a learning-based or behavioral model of user feedback for propensity estimation \cite{joachims2017unbiased,yang2018unbiased}, which remains the central concern of this approach.
 While many works rely on a propensity matrix that is assumed to be known, e.g.
@@
 
-The IPS method in our context is, essentially, a method of weighting training examples to correct the bias in the training data.
-Equivalent approaches such as importance weighting are widely used for domain adaptation in fields other than recommender systems \cite{sugiyama2007covariate} \cite{zhang2018importance}.
-Discussions on the IPS-based domain adaptation for countering selection bias, which is most relevant to recommender systems, remain limited due to the hardness of evaluating propensities.
+The IPS method in our context is essentially a method of weighting training examples to correct the bias in the training data.
+Equivalent approaches such as importance weighting are widely used for domain adaptation in fields other than recommender systems \cite{sugiyama2007covariate,zhang2018importance}.
+Discussions on countering selection bias, e.g. in recommender systems, with IPS-based domain adaptation remain limited due to the hardness of evaluating propensities.
+% Discussions on the IPS-based domain adaptation for countering selection bias, which is most relevant to recommender systems, remain limited due to the hardness of evaluating propensities.
 
@@
 
-\subsection{Simulated Environment for Controlling Bias}
+\subsection{Simulated Environment for Controlling Bias \textcolor{blue}{and Quantization}}
 Since selection bias is uncontrollable in a dataset completely drawn from real-world, we have to adopt a simulated environment \cite{krauth2020offline} with both semi-synthetic and synthetic datasets explained below.
 In our environment, we propose the \textbf{softmax observation model} and introduce a hyperparameter $\beta$ to control the degree of bias.
-For a rating matrix $R$, the corresponding probability matrix of each rating being observed is $\Pr(R_{u,i}\text{ is observed}) = k\text{softmax}(\beta R_{u,i})$, where $k$ is pre-set so that the expected proportion of observed ratings is controlled.
+For a rating matrix $R$, the corresponding probability matrix of each rating being observed is $\Pr(R_{u,i}\text{ is observed}) = k\text{softmax}(\beta R_{u,i})$, where $k$ is set so that the expected proportion of observed ratings is controlled.
 In this observation model, a larger $\beta$ induces more significant bias at an exponential rate.
@@
 
+\textcolor{blue}{Mention how we control quantization: mapping the same underlying continuous-valued $R_{ui}$ to $n$ different quantities evenly spaced between $0$ and $1$, i.e.
+0 and 1 or 0, 0.25, 0.5, 0.75, 1}
+
 \subsection{Datasets}
 \textbf{Imputed ML100K Dataset.}
-The ML100K dataset provides 100 thousand MNAR ratings across 1683 movies rated by 944 users and is the standard large-scale dataset used for recommender systems.
-Since we need the ground truth ratings for controlling bias, we impute the missing ratings using standard matrix factorization \cite{krauth2020offline}.
-Ratings are normalized between 0 and 1 for consistency with the other dataset.
+The ML100K dataset provides 100 thousand MNAR ratings across 1683 movies rated by 944 users and is a standard dataset used for recommender systems.
+Since we need the ground truth ratings for controlling bias, we use imputed ratings computed via standard matrix factorization \cite{krauth2020offline}.
+We normalize ratings between 0 and 1 for consistency with the other dataset.
 
 \textbf{Latent Factors Simulated Dataset.}
-The latent factors dataset is a synthetic dataset that models real-world user behavior.
-Users and item both have Gaussian latent vectors to simulate preferences \cite{koren2008factorization} \cite{koren2009matrix}; both also have biases.
+The latent factors dataset is a synthetic dataset that models users and item with Gaussian latent vectors to simulate preferences \cite{koren2008factorization,koren2009matrix}; both also have biases.
 The environment is provided by \cite{krauth2020offline}.
@@
 \end{figure}
-Figure 1 shows our results on the latent factors dataset with 3 classic algorithms: user-KNN, item-KNN, and SVD matrix factorization.
+Figure 1 shows the performance on the latent factors dataset of 3 classic algorithms: user-KNN, item-KNN, and SVD matrix factorization.
+\textcolor{blue}{Performance is evaluated by RMSE on the dense quantized (or not?) rating matrix}.
 The three differently quantized variants come from the same source distribution (the dataset).
-Gaussian noise is introduced for simulating real-world user noise.
-Shaded regions denote the regions between quartiles.
+Gaussian noise is introduced for simulating real-world user variability.
+Shaded regions denote the regions between quartiles \textcolor{blue}{overs [number] trials of re-sampled training data}.
 
-Although the RMSE grows consistently as $\beta$ increases, differently quantized datasets do not exhibit significantly different growth rates.
+Although the RMSE grows consistently as sampling bias $\beta$ increases, differently quantized datasets do not exhibit significantly different growth rates.
 The discrepancy between the RMSE of differently quantized datasets with the same $\beta$ can be ascribed to the inherent information loss from quantization \cite{widrow1996statistical}.
@@
 We thus conclude it is a viable approach to use the more quantized, less biased dataset for propensity estimation.
-Furthermore, figure 1.
-visualizes the damaging effect of selection bias.
-The less quantized datasets breaks even with the most coarse binary dataset at $\beta=2.5$ versus $\beta=0$, suggesting the practical importance of debiasing recommender systems.
-The above experiments are repeated under a semi-synthetic environment and the conclusion holds.
+Furthermore, Figure 1 visualizes the damaging effect of selection bias.
+The less quantized data with high sampling bias ($\beta=2.5$) performs just as poorly as the coarse binary dataset with no sampling bias ($\beta=0$),
+% The performance of the less quantized datasets breaks even with the most  at $\beta=2.5$ versus $\beta=0$, 
+suggesting the practical importance of debiasing recommender systems. 
+The above experiments are repeated with the semi-synthetic ML100K environment and the conclusion holds.
 
 \section{Cross-Dataset Propensity Estimation}
-With the above guarantee, we now turn to our cross-dataset matrix factorization model.
+With the above result, we now turn to our cross-dataset matrix factorization model.
 Matrix factorization is a simple recommender model that decomposes the rating matrix based on known entries and then predicts the unknown entries \cite{koren2009matrix}.
@@
 \end{equation}
-where $A = \{ b_u, b_i, \mu \}$ represents the standard bias parameters (offset), $V$ and $W$ are the decomposed vectors, $\hat{Y} = V_u^TW_i + A$ is the predicted rating, $N$ is the number of ratings, and $c \Vert A \Vert^2$ is the regularizer.
+where $A = \{ b_u, b_i, \mu \}$ represents the standard bias parameters (offset), $V$ and $W$ are the decomposed vectors, $\hat{Y} = V_u^TW_i + A$ is the predicted rating, $N$ is the number of observed ratings, and $c \Vert A \Vert^2$ is the regularizer.
 The inverse propensity scores $\nicefrac{1}{P_{u,i}}$ are multiplied to each rating during learning, which is analogous to re-weighting ratings based on their biases.
@@
 The numerator can be easily approximated from the dataset, but the denominator requires additional data from less biased sources.
-In this paper, we estimates the denominator from the less biased dataset using a categorical naive-bayes model.
+In this paper, we estimate the denominator from the less biased dataset using a categorical naive-bayes model.
 We use $D'$ as a mask to hide biased ratings in the training set from propensity estimation.
@@
 Empirically, our framework will work as long as the highly quantized dataset is less biased.
-We selected the root mean square error (RMSE) and the mean absolute error (MAE) as performance metrics.
+We selected the root mean square error (RMSE) and the mean absolute error (MAE) \textcolor{blue}{over the dense rating matrix (?)} as performance metrics.
 
@@
 \textbf{Matrix Factorization (MF.)}
-As a simple baseline, we adopt the standard matrix factorization that does not adopt propensity estimation or importance weighting.
+As a simple baseline, we adopt the standard matrix factorization that does not adopt propensity estimation or importance weighting \textcolor{blue}{and uses only the training set}.
 
@@
 \section{Conclusion}
-We propose an effective and data-efficient method of combing differently sourced datasets for training recommender systems.
+We propose an effective and data-efficient method of combining differently sourced datasets for training recommender systems.
 Our approach provides yet another way of accurate propensity estimation and explores a new potential of cross-dataset joint learning.
 We demonstrate that our method outperforms naive propensity estimation and straightforward dataset mixing.
-and  We believe real-world recommender systems will benefit from data-centric frameworks like ours.
+We believe real-world recommender systems will benefit from data-centric frameworks like ours.
 

### 2022-09-26 paper/paper.tex
Final submitted version

@@
 
-\subsection{Simulated Environment for Controlling Bias \textcolor{blue}{and Quantization}}
+\subsection{Simulated Environment for Controlling Bias and Quantization}
 Since selection bias is uncontrollable in a dataset completely drawn from real-world, we have to adopt a simulated environment \cite{krauth2020offline} with both semi-synthetic and synthetic datasets explained below.
 In our environment, we propose the \textbf{softmax observation model} and introduce a hyperparameter $\beta$ to control the degree of bias.
-For a rating matrix $R$, the corresponding probability matrix of each rating being observed is $\Pr(R_{u,i}\text{ is observed}) = k\text{softmax}(\beta R_{u,i})$, where $k$ is set so that the expected proportion of observed ratings is controlled.
+For a rating matrix $R$, the corresponding probability matrix of each rating being observed is $\Pr(R_{u,i}\text{ is observed}) = k\,\text{softmax}(\beta R_{u,i})$, where $k$ is set so that the expected proportion of observed ratings is controlled.
 In this observation model, a larger $\beta$ induces more significant bias at an exponential rate.
 In our experiment we assume constantly $10\%$ of ratings are observed among $1000$ users and $1000$ items.
+Quantization is simulated by mapping the same underlying continuous-valued rating matrix $R$ to $n$ different quantities evenly spaced between $0$ and $1$.
+In all experiments we used only 2, 3, and 5 as quantization levels.
 % \begin{figure}
@@
 % \end{figure}
-
-\textcolor{blue}{Mention how we control quantization: mapping the same underlying continuous-valued $R_{ui}$ to $n$ different quantities evenly spaced between $0$ and $1$, i.e.
-0 and 1 or 0, 0.25, 0.5, 0.75, 1}
 
@@
 Figure 1 shows the performance on the latent factors dataset of 3 classic algorithms: user-KNN, item-KNN, and SVD matrix factorization.
-\textcolor{blue}{Performance is evaluated by RMSE on the dense quantized (or not?) rating matrix}.
+Performance is evaluated by RMSE on the dense ground truth rating matrix.
 The three differently quantized variants come from the same source distribution (the dataset).
 Gaussian noise is introduced for simulating real-world user variability.
-Shaded regions denote the regions between quartiles \textcolor{blue}{overs [number] trials of re-sampled training data}.
+Shaded regions denote the regions between quartiles overs 10 trials of re-sampled training data.
 
@@
 Empirically, our framework will work as long as the highly quantized dataset is less biased.
-We selected the root mean square error (RMSE) and the mean absolute error (MAE) \textcolor{blue}{over the dense rating matrix (?)} as performance metrics.
+We selected the root mean square error (RMSE) and the mean absolute error (MAE) over the dense ground truth rating matrix as performance metrics.
 
@@
 \textbf{Matrix Factorization (MF.)}
-As a simple baseline, we adopt the standard matrix factorization that does not adopt propensity estimation or importance weighting \textcolor{blue}{and uses only the training set}.
+As a simple baseline, we adopt the standard matrix factorization that does not adopt propensity estimation or importance weighting and uses only the less quantized training set.
 

### 2022-11-16 paper/paper.tex
Edit acknoledgement
Final version

@@
 % to compile a camera-ready version, add the [final] option, e.g.:
-\usepackage[]{neurips_2022}
+\usepackage[final]{neurips_2022}
 \bibliographystyle{plain}
@@
 \textbf{Imputed ML100K Dataset.}
-The ML100K dataset provides 100 thousand MNAR ratings across 1683 movies rated by 944 users and is a standard dataset used for recommender systems.
+The ML100K dataset provides 100 thousand MNAR (missing-not-at-random) ratings across 1683 movies rated by 944 users and is a standard dataset used for recommender systems.
 Since we need the ground truth ratings for controlling bias, we use imputed ratings computed via standard matrix factorization \cite{krauth2020offline}.
@@
 \end{equation}
-where $A = \{ b_u, b_i, \mu \}$ represents the standard bias parameters (offset), $V$ and $W$ are the decomposed vectors, $\hat{Y} = V_u^TW_i + A$ is the predicted rating, $N$ is the number of observed ratings, and $c \Vert A \Vert^2$ is the regularizer.
+where $A = \{ b_u, b_i, \mu \}$ represents the standard bias parameters (offset), $Y$ is the ground truth ratings, $V$ and $W$ are the decomposed vectors, $\hat{Y} = V_u^TW_i + A$ is the predicted rating, $N$ is the number of observed ratings, and $c \Vert A \Vert^2$ is the regularizer.
 The inverse propensity scores $\nicefrac{1}{P_{u,i}}$ are multiplied to each rating during learning, which is analogous to re-weighting ratings based on their biases.
@@
 \begin{equation} 
-  P_{u,i} = \Pr(Y_{u,i} \text{ is observed} \mid Y_{u,i} = r_{u,i}) = \frac{\Pr(Y_{u,i} = r_{u,i} \mid Y_{u,i} \text{ is observed})}{\Pr(Y_{u,i} = r_{u,i})},
+  P_{u,i} = \Pr(Y_{u,i} \text{ is observed} \mid Y_{u,i} = r_{u,i}) = \frac{\Pr(Y_{u,i} = r_{u,i} \mid Y_{u,i} \text{ is observed})\Pr( Y_{u,i} \text{ is observed})}{\Pr(Y_{u,i} = r_{u,i})},
 \end{equation}
@@
 
-% \section*{Acknowledgements}
-% This work is supported in part by a Cornell CSURP Grant,  {\color{blue}{\textit{is there some source of funding you'd like to acknowledge?}}}
+\section*{Acknowledgements}
+This work is supported in part by a Cornell CSURP Grant and by a research gift from Wayfair.
 
